---
title: "Comparison plots"
author: "Ludger Goeminne"
date: "04/07/2019"
output: html_document
---

# Figures in the main article

This document will allow you to reproduce all figures in the main article.

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
source(paste0(wd,"/R/plot_functions.R"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

Important: make sure you have run the CPTAC and HEART analyses, or load the necessary files as follows:

```{r}
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC2.RData"))
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC_RegEM_imp.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_RegEM_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_co_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_Hurdle.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_mixture.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_ProDA_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_DEP.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_DEP_noimp.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_MSstats_noimp_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_MSstats_full.RData"))
load(paste0(wd, "/save_files_CPTAC/CPTAC_Perseus_full.RData"))
load(paste0(wd, "/save_files_CPTAC/CPTAC_Perseus_imp_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_ProPCA_full_QN_int.RData"))
load(paste0(wd, "/save_files_PXD006675/peptides_HEART2.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_Hurdle.RData"))
load(paste0(wd, "/save_files_PXD006675/overlap_PHM.RData"))
```

## 1. Evaluate the imputation

### 1.1. Evaluate Perseus imputation (Fig. 1 a)

We make a histogram to evaluate the imputation. Blue are the original values, red are the imputed values. For the sample-wise effect of imputation, see supp_comparison_plots.Rmd.

```{r eval = TRUE}
cex.main <- 1.2
cex.lab <- 1
cex <- 1
lwd <- 1
# #Plotting this plot in SVG together with the next one:
# grDevices::svg(paste0(wd,"/missing_imp.svg"), width=20, height=10)
# cex.main <- 3.5
# cex.lab <- 3
# cex <- 2.5
# par(mfrow=c(1,2))
# par(mar=c(5.1*cex,4.1*cex+3,4.1*cex,2.1*cex+7))

# Histogram for the effect of Perseus imputation on sample A5 (Fig. 1 a)

# Import non-imputed data
file.proteinGroups <- paste0(wd,"/datasets/CPTAC/proteinGroups_preprocessed.txt")
exprs_col <- grepEcols(file.proteinGroups, "LFQ.intensity.", split = "\t")
proteinGroups.CPTAC <- readMSnSet2(file.proteinGroups, ecol = exprs_col, fnames = c("Protein.IDs"), sep = "\t")
sampleNames(proteinGroups.CPTAC) <- str_replace(sampleNames(proteinGroups.CPTAC), "LFQ.intensity.", "") %>% make.names

# Import Perseus-imputed data
file.proteinGroups <- paste0(wd,"/datasets/CPTAC/proteinGroups_preprocessed_imputed.txt")
exprs_col <- grepEcols(file.proteinGroups, "LFQ.intensity.", split = "\t")
proteinGroups.CPTAC.Perseus.imp <- readMSnSet2(file.proteinGroups, ecol = exprs_col, fnames = c("Protein.IDs"), sep = "\t")
sampleNames(proteinGroups.CPTAC.Perseus.imp) <- str_replace(sampleNames(proteinGroups.CPTAC.Perseus.imp), "LFQ.intensity.", "") %>% make.names

sample.names <- paste0(sampleNames(proteinGroups.CPTAC) %>% substr(3, 3), sampleNames(proteinGroups.CPTAC) %>% substr(5, 5))

# Histogram for the effect of Perseus imputation on protein level in sample A5
i <- 5
p1 <- hist(exprs(proteinGroups.CPTAC.Perseus.imp[,i]), breaks=25, plot = FALSE)
p2 <- hist(exprs(proteinGroups.CPTAC[,i]), breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(12,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("Protein-level Perseus imputation sample ", sample.names[i]), xlab = "log2(LFQ intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(12,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram

length(is.na(exprs(proteinGroups.CPTAC[,i])))-sum(is.na(exprs(proteinGroups.CPTAC[,i])))
# 978 observed LFQ intensities
sum(is.na(exprs(proteinGroups.CPTAC[,i])))
# 484 imputed values

```

### 1.2. Evaluate RegEM imputation (Fig. 1 b)

We make a histogram to evaluate the imputation. Blue are the original values, red are the imputed values. For the sample-wise effect of imputation, see supp_comparison_plots.Rmd.

```{r eval = TRUE}

# Histogram for the effect of RegEM imputation on protein level in sample A5 (Fig. 1 b)

sample.names <- paste0(sampleNames(peptides.CPTAC2) %>% substr(3, 3), sampleNames(peptides.CPTAC2) %>% substr(5, 5))

i <- 5
p1 <- hist(exprs(peptides.CPTAC.RegEM.imp[,i]), breaks=25, plot = FALSE)
p2 <- hist(exprs(peptides.CPTAC2[,i]), breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(12,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("Peptide-level RegEM imputation sample ", sample.names[i]), xlab = "log2(peptide intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(12,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram

length(is.na(exprs(peptides.CPTAC2[,i])))-sum(is.na(exprs(peptides.CPTAC2[,i])))
# 6766 observed peptide intensities
sum(is.na(exprs(peptides.CPTAC2[,i])))
# 2611 imputed values

# par(mfrow=c(1,1))
# par(mar=c(5.1,4.1,4.1,2.1))
# 
# # End SVG plot
# dev.off()
```

## 2. FDR-FTP plots

### 2.1. Prepare the FDR-FTP plots

```{r}
# Due to different preprocessing methods, some proteins cannot be detected with some methods.

# By default we use the full set, i.e. all protein identifiers over all methods.
# Note that all methods used either the preprocessing of MSqRob, of Perseus o of MSstats.
all.proteins <- unique(unlist(lapply(list(res.CPTAC.Hurdle, CPTAC.Perseus.full, res.CPTAC.MSstats.noimp.full), "[", "protein")))
full.Base <- tibble(protein = rep(all.proteins, 3), 
                    UPS = grepl("ups", protein),
                    contrast = rep(c("condition6B-condition6A",
                  "condition6C-condition6B",
                  "condition6C-condition6A"), each = length(all.proteins)))
```

### 2.2. Make the introductory FDR-FTP plots (Fig. 2)

```{r eval = TRUE}
res.list <- list(res.CPTAC.MSstats.full[!is.infinite(res.CPTAC.MSstats.full$log2FC),],
                 res.CPTAC.MSstats.noimp.full[!is.infinite(res.CPTAC.MSstats.noimp.full$log2FC),],
                 CPTAC.Perseus.imp.full,
                 CPTAC.Perseus.full,
                 res.CPTAC.co.full,
                 res.CPTAC.RegEM.full,
                 res.CPTAC.full)

res.list <- lapply(res.list, function(x){
    left_join(full.Base, x)
}
  )

contrast.vec <- c("condition6B-condition6A",
                  "condition6C-condition6A",
                  "condition6C-condition6B")

names(contrast.vec) <- c("Condition 6B vs. 6A",
                  "Condition 6C vs. 6A",
                  "Condition 6C vs. 6B")

colors <- c("#808000", "green", "#42B7BA", "#1B2944", "#E15E9E", "black", "#FF681E") # darkblue: "#1B2944"  gray: "#50FF00"

sort.list <- list(c("adj.pvalue", "pvalue", "Tvalue", "log2FC"),
                  c("adj.pvalue", "pvalue", "Tvalue", "log2FC"),
                  c("q.value", "p.value", "Test.statistic", "Difference"),
                  c("q.value", "p.value", "Test.statistic", "Difference"),
                  c("qvalue", "pvalue", "t", "logOR"),
                  c("qvalue", "pvalue", "t", "logFC"),
                  c("qvalue", "pvalue", "t", "logFC"))

PlotFDPTPR(res.list, contrast.vec, colors, sort.list, TPcol = c("UPS"), plotSVG = FALSE) # TRUE
```

## 3. Make the FDR-FTP plots for MaxLFQ-based state-of-the-art methods (Fig. 3)

```{r}
res.list <- list(CPTAC.Perseus.imp.full,
     CPTAC.Perseus.full,
     res.CPTAC.DEP,
     res.CPTAC.mixture,
     res.CPTAC.ProDA.full,
     res.CPTAC.DEP.noimp,
     res.CPTAC.co.full,
     res.CPTAC.full,
     res.CPTAC.Hurdle)

res.list <- lapply(res.list, function(x){
    left_join(full.Base, x)
}
  )

contrast.vec <- c("condition6B-condition6A",
                  "condition6C-condition6A",
                  "condition6C-condition6B")

names(contrast.vec) <- c("Condition 6B vs. 6A",
                  "Condition 6C vs. 6A",
                  "Condition 6C vs. 6B")

colors <- c("#42B7BA", "#1B2944", "peru", "#FFC30B", "#787878", "yellow3", "#E15E9E", "#FF681E", "#5A2A82")

sort.list <- list(c("q.value", "p.value", "Test.statistic", "Difference"),
     c("q.value", "p.value", "Test.statistic", "Difference"),
     c("qvalue", "pvalue", "logFC"),
     c("qvalue", "pvalue"),
     c("qval", "pval", "t", "logFC"),
     c("qvalue", "pvalue", "logFC"),
     c("qvalue", "pvalue", "t", "logOR"),
     c("qvalue", "pvalue", "t", "logFC"),
     c("qchisq", "pchisq", "chisq", NA))

PlotFDPTPR(res.list, contrast.vec, colors, sort.list, TPcol = c("UPS"), plotSVG = FALSE) # TRUE
```

## 4. Make the FDR-FTP plots for peptide-based state-of-the-art methods methods (Fig. 4)

```{r eval = TRUE}
res.list <- list(res.CPTAC.MSstats.full[!is.infinite(res.CPTAC.MSstats.full$log2FC),],
                 res.CPTAC.MSstats.noimp.full[!is.infinite(res.CPTAC.MSstats.noimp.full$log2FC),],
                 res.CPTAC.ProPCA.full.QN.int, 
                 res.CPTAC.RegEM.full,
                 res.CPTAC.co.full,
                 res.CPTAC.full,
                 res.CPTAC.Hurdle)

res.list <- lapply(res.list, function(x){
    left_join(full.Base, x)
}
  )

contrast.vec <- c("condition6B-condition6A",
                  "condition6C-condition6A",
                  "condition6C-condition6B")

names(contrast.vec) <- c("Condition 6B vs. 6A",
                  "Condition 6C vs. 6A",
                  "Condition 6C vs. 6B")

colors <- c("#808000", "green", "red2", "black", "#E15E9E", "#FF681E", "#5A2A82") # darkblue: "#1B2944"  gray: "#50FF00"

sort.list <- list(c("adj.pvalue", "pvalue", "Tvalue", "log2FC"),
                  c("adj.pvalue", "pvalue", "Tvalue", "log2FC"),
                  c("adj.P.Val", "P.Value", "t", "logFC"), 
                  c("qvalue", "pvalue", "t", "logFC"),
                  c("qvalue", "pvalue", "t", "logOR"),
                  c("qvalue", "pvalue", "t", "logFC"),
                  c("qchisq", "pchisq", "chisq", NA))

PlotFDPTPR(res.list, contrast.vec, colors, sort.list, TPcol = c("UPS"), plotSVG = FALSE) # TRUE
```

## 5. Figures for the HEART dataset (Fig. 5)

### 5.1. Generate the data for the Venn diagrams for the HEART dataset (Fig. 5 a)

For the plots with the overlap for the top 500, top 1000 and all significant genes, see supp_comparison_plots.Rmd.

```{r eval = TRUE}
### Venn diagram atrial vs. ventricular region ###

### First 1500 most significant genes ###

# A. Select top 1500 DA gene identifiers for the Perseus analysis of ventricles vs. atria
genes.Perseus.AvsV.1500 <- (Perseus.AvsV.ov %>% pull(`Gene names`) %>% unique)[1:1500]

# B. Select top 1500 DA gene identifiers for the Hurdle analysis of atria vs. ventricles
genes.Hurdle.AvsV.1500 <- hurdle.AvsV.ov[1:1500,]$gene.name

# C. Select top 1500 DA gene identifiers for the MSqRob analysis of atria vs. ventricles
genes.MSqRob.AvsV.1500 <- MSqRob.AvsV.ov[1:1500,]$gene.name

# MSqRob alone

sum(!(genes.MSqRob.AvsV.1500 %in% genes.Hurdle.AvsV.1500) & !(genes.MSqRob.AvsV.1500 %in% genes.Perseus.AvsV.1500))
# 247

# Hurdle alone

sum(!(genes.Hurdle.AvsV.1500 %in% genes.MSqRob.AvsV.1500) & !(genes.Hurdle.AvsV.1500 %in% genes.Perseus.AvsV.1500))
# 164

# Perseus alone

sum(!(genes.Perseus.AvsV.1500 %in% genes.MSqRob.AvsV.1500) & !(genes.Perseus.AvsV.1500 %in% genes.Hurdle.AvsV.1500))
# 596

# Overlap MSqRob and Hurdle

sum((genes.MSqRob.AvsV.1500 %in% genes.Hurdle.AvsV.1500) & !(genes.MSqRob.AvsV.1500 %in% genes.Perseus.AvsV.1500))
# 507

# Overlap MSqRob and Perseus

sum((genes.MSqRob.AvsV.1500 %in% genes.Perseus.AvsV.1500) & !(genes.MSqRob.AvsV.1500 %in% genes.Hurdle.AvsV.1500))
# 75

# Overlap Hurdle and Perseus

sum((genes.Hurdle.AvsV.1500 %in% genes.Perseus.AvsV.1500) & !(genes.Hurdle.AvsV.1500 %in% genes.MSqRob.AvsV.1500))
# 158

# Overlap everything

sum((genes.MSqRob.AvsV.1500 %in% genes.Perseus.AvsV.1500) & (genes.MSqRob.AvsV.1500 %in% genes.Hurdle.AvsV.1500))
# 671

# Control:
length(genes.MSqRob.AvsV.1500)
247+507+75+671

length(genes.Hurdle.AvsV.1500)
164+507+158+671

length(genes.Perseus.AvsV.1500)
596+158+75+671
```

### 5.2. Make scatter plot for the HEART dataset (Fig. 5 b)

```{r eval = TRUE}
### Make the plots ###
  
cex.main <- 1.2
cex.lab <- 1
cex <- 1
lwd <- 1
# grDevices::svg("HEART_full_scatter.svg", width=20, height=10)
# cex.main <- 4.3
# cex.lab <- 3.3
# cex <- 3
# lwd <- 6

HEART_A_vsV <- res.HEART.Hurdle %>% filter(contrast == "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3") %>% na.omit %>% arrange(pchisq %>% desc)

cols <-  colorRampPalette(c("#FF3100", "#FCFF00", "#45FE4F",
                            "#00FEFF", "#000099" 
                            ))(256)

plot(HEART_A_vsV$logFC, HEART_A_vsV$logOR, col = cols[ceiling(rank(HEART_A_vsV$pchisq)/length(HEART_A_vsV$pchisq)*256)], pch = 20, xlab = "MSqRob log2FC", ylab = "Binomial log2OR", cex = cex*0.5, cex.lab = cex.lab, cex.axis = cex, cex.main = cex.main, las = 1, frame.plot = FALSE, main = "Atrial vs. ventricular regions") #[1:1000]
abline(v = 0, col = "black", lwd = cex)
abline(h = 0, col = "black", lwd = cex)

# dev.off()

### Colors for 1%, 5% and 10% FDR ###

# Color for 1% FDR
which(HEART_A_vsV$qchisq < 0.01)[1]
# 5435
cols[ceiling(rank(HEART_A_vsV$pchisq)[5435]/length(HEART_A_vsV$pchisq)*256)]
cols[ceiling(rank(HEART_A_vsV$pchisq)[5434]/length(HEART_A_vsV$pchisq)*256)]
# "#FDB500"

# Color for 5% FDR
which(HEART_A_vsV$qchisq < 0.05)[1]
# 4778
cols[ceiling(rank(HEART_A_vsV$pchisq)[4778]/length(HEART_A_vsV$pchisq)*256)]
cols[ceiling(rank(HEART_A_vsV$pchisq)[4777]/length(HEART_A_vsV$pchisq)*256)]
# "#F2FE04"

# Color for 10% FDR
which(HEART_A_vsV$qchisq < 0.1)[1]
# 4290
cols[ceiling(rank(HEART_A_vsV$pchisq)[4290]/length(HEART_A_vsV$pchisq)*256)]
cols[ceiling(rank(HEART_A_vsV$pchisq)[4289]/length(HEART_A_vsV$pchisq)*256)]
# "#BCFE1B"
```