---
title: "Supplementary figures"
author: "Ludger Goeminne"
date: "09/07/2018"
output: html_document
---

# Figures in supplementary material

This document will allow you to reproduce all supplementary figures.

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
source(paste0(wd,"/R/plot_functions.R"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

Important: make sure you have run the CPTAC and HEART analyses, or load the necessary files as follows:

```{r}
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC2.RData"))
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC_KNN1.RData"))
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC_Perseus_imp.RData"))
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC_mixed.RData"))
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC_RegEM_imp.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_co_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_sc_co_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_edgeR_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_edgeR_sc_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_beta_binom_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_beta_binom_sc_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_Hurdle.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_PI_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_KNN1_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_RegEM_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_mixed_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_ProDA_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_ProPCA_full_QN_int.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_mixture.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_DEP.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_DEP_noimp.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_MSstats_full.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_MSstats_noimp_full.RData"))
load(paste0(wd, "/save_files_CPTAC/CPTAC_Perseus_full.RData"))
load(paste0(wd, "/save_files_CPTAC/CPTAC_Perseus_imp_full.RData"))
load(paste0(wd, "/save_files_PXD006675/peptides_HEART2.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_Hurdle.RData"))
load(paste0(wd, "/save_files_PXD006675/overlap_PHM.RData"))
```

## 1. Prepare for imputation figures: imputation at the protein level

```{r eval=TRUE}
# Import non-imputed data
file.proteinGroups <- paste0(wd,"/datasets/CPTAC/proteinGroups_preprocessed.txt")
exprs_col <- grepEcols(file.proteinGroups, "LFQ.intensity.", split = "\t")
proteinGroups.CPTAC <- readMSnSet2(file.proteinGroups, ecol = exprs_col, fnames = c("Protein.IDs"), sep = "\t")
sampleNames(proteinGroups.CPTAC) <- str_replace(sampleNames(proteinGroups.CPTAC), "LFQ.intensity.", "") %>% make.names

# Import Perseus-imputed data
file.proteinGroups <- paste0(wd,"/datasets/CPTAC/proteinGroups_preprocessed_imputed.txt")
exprs_col <- grepEcols(file.proteinGroups, "LFQ.intensity.", split = "\t")
proteinGroups.CPTAC.Perseus.imp <- readMSnSet2(file.proteinGroups, ecol = exprs_col, fnames = c("Protein.IDs"), sep = "\t")
sampleNames(proteinGroups.CPTAC.Perseus.imp) <- str_replace(sampleNames(proteinGroups.CPTAC.Perseus.imp), "LFQ.intensity.", "") %>% make.names

sample.names <- paste0(sampleNames(proteinGroups.CPTAC) %>% substr(3, 3), sampleNames(proteinGroups.CPTAC) %>% substr(5, 5))

### Impute with KNN-1
proteinGroups.CPTAC.KNN1 <- impute(proteinGroups.CPTAC, method = "knn")

cex.main <- 1.2
cex.lab <- 1
cex <- 1
par(mfrow=c(1,1))

### Impute with mixed imputation
count.proteins <- MSnSet2df(proteinGroups.CPTAC) %>% as_tibble %>% group_by(feature, condition) %>% summarize(n())
proteins_MNAR <- names(which(table(count.proteins$feature) != 3))

## Get a logical vector
  MNAR <- rownames(exprs(proteinGroups.CPTAC)) %in% proteins_MNAR
  print('Number of proteins that are Missing Not At Random')
  print(table(MNAR))

 ## Perform a mixed imputation
  # The DEP example in the vignette considers a protein to have missing values not at random (MNAR) if it has missing values in all replicates of at least one condition.
randna <- !MNAR
proteinGroups.CPTAC.mixed <- proteinGroups.CPTAC

exprs(proteinGroups.CPTAC.mixed)[randna, ] <- exprs(impute(proteinGroups.CPTAC, "knn"))[randna,]
exprs(proteinGroups.CPTAC.mixed)[!randna, ] <- exprs(impute(proteinGroups.CPTAC, "MinProb"))[!randna,]

### Impute with RegEM imputation
# 1. First remove this one protein that has no observed values (RegEM cannot handle this)
proteinGroups.CPTAC2 <- proteinGroups.CPTAC[rowSums(is.na(Biobase::exprs(proteinGroups.CPTAC))) != ncol(Biobase::exprs(proteinGroups.CPTAC)),]

# 2. Write protein info as tab delimited file
write.table(cbind(protein = row.names(Biobase::exprs(proteinGroups.CPTAC2)), Biobase::exprs(proteinGroups.CPTAC2)), file = paste0(wd,"/datasets/CPTAC/proteinGroups_CPTAC.txt"), sep = "\t", quote = FALSE, row.names = FALSE, col.names = TRUE)

# 3. Download the MatLab files from https://github.com/tapios/RegEM#installation.

# 4. Execute the following code in MatLab:
# X=readmatrix("proteinGroups_CPTAC.txt")
# X_imputed=regem(X)
# writematrix(X_imputed,"proteinGroups_CPTAC_RegEM_imp.txt",'Delimiter','tab')

# The resulting proteinGroups_CPTAC_RegEM_imp.txt file can be found in the "/datasets/CPTAC" folder.

# 5. Import the imputed values
exprs.RegEM.imp <- read.table(file = paste0(wd,"/datasets/CPTAC/proteinGroups_CPTAC_RegEM_imp.txt"), sep = "\t", header = FALSE, quote = "", dec = ".")

# 6. Preprocess the data frame
exprs.RegEM.imp <- exprs.RegEM.imp[,-1]
dimnames(exprs.RegEM.imp) <- dimnames(exprs(proteinGroups.CPTAC2))

# 7. Put the imputed values in the MSnSet object proteins.CPTAC.RegEM.imp
proteinGroups.CPTAC.RegEM.imp <- proteinGroups.CPTAC2
Biobase::exprs(proteinGroups.CPTAC.RegEM.imp) <- exprs.RegEM.imp[,1:27] %>% as.matrix
```

## 2. Evaluate Perseus imputation: protein level (Supplementary Fig. 2 and 3)

We here show the sample-wise effect of Perseus imputation at the level of MaxQuant's LFQ-values (protein level). Blue are the original values, red are the imputed values.

```{r eval=TRUE}
# 1. Histograms for the effect of Perseus imputation LFQ-level (Supplementary Fig. 2)

# grDevices::svg(paste0(wd,"/Perseus_protein_sample_imp.svg"), width=60, height=90)
# cex.main <- 10
# cex.lab <- 8
# cex <- 6
# par(mfrow=c(1,3))
# par(mar=c(5.1*cex/2,4.1*cex/2,4.1*cex/2,2.1*cex/2))
# par(mfrow=c(9,3))

for(i in 1:ncol(exprs(proteinGroups.CPTAC.Perseus.imp))){
p1 <- hist(exprs(proteinGroups.CPTAC.Perseus.imp[,i]), breaks=25, plot = FALSE)
p2 <- hist(exprs(proteinGroups.CPTAC[,i]), breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(12,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("Perseus imputation sample ", sample.names[i]), xlab = "log2(MaxLFQ intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(12,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
}

# dev.off()
```

## 3. Evaluate Perseus imputation: peptide level (Supplementary Fig. 3 and 4)

We here show the sample-wise effect of Perseus imputation at the peptide level. Blue are the original values, red are the imputed values.

```{r eval=TRUE}

# 2. Histograms for the effect of Perseus imputation at peptide-level (Supplementary Fig. 3)

# grDevices::svg(paste0(wd,"/Perseus_sample_imp.svg"), width=60, height=90)
# cex.main <- 10
# cex.lab <- 8
# cex <- 6
# par(mfrow=c(1,3))
# par(mar=c(5.1*cex/2,4.1*cex/2,4.1*cex/2,2.1*cex/2))
# par(mfrow=c(9,3))

for(i in 1:ncol(exprs(peptides.CPTAC.Perseus.imp))){
p1 <- hist(exprs(peptides.CPTAC.Perseus.imp[,i]), breaks=25, plot = FALSE)
p2 <- hist(exprs(peptides.CPTAC2[,i]), breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(12,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("Perseus imputation sample ", sample.names[i]), xlab = "log2(peptide intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(12,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
}

# dev.off()

```

## 4. Histograms for the effects of all imputations at protein level, grouped per condition (Supplementary Fig. 5)

```{r eval=TRUE}

condition.names <- unique(sampleNames(proteinGroups.CPTAC) %>% substr(3, 3))

cex.main <- 1.2
cex.lab <- 1
cex <- 1
par(mfrow=c(1,1))

# grDevices::svg(paste0(wd,"/histograms_all_imputations_proteins.svg"), width=60, height=120)
# cex.main <- 10
# cex.lab <- 8
# cex <- 6
# par(mfrow=c(4,3))
# par(mar=c(5.1*cex/2,4.1*cex/2,4.1*cex/2,2.1*cex/2))
# par(mfrow=c(9,3))

# Perseus imputation
for(i in 1:3){
p1 <- hist(exprs(proteinGroups.CPTAC.Perseus.imp)[,which(grepl(condition.names[i], sampleNames(proteinGroups.CPTAC.Perseus.imp)))], breaks=25, plot = FALSE)
p2 <- hist(exprs(proteinGroups.CPTAC)[,which(grepl(condition.names[i], sampleNames(proteinGroups.CPTAC)))], breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(13,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("Perseus imputation condition ", condition.names[i]), xlab = "log2(MaxLFQ intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(13,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
}

# kNN imputation
for(i in 1:3){
p1 <- hist(exprs(proteinGroups.CPTAC.KNN1)[,which(grepl(condition.names[i], sampleNames(proteinGroups.CPTAC.KNN1)))], breaks=25, plot = FALSE)
p2 <- hist(exprs(proteinGroups.CPTAC)[,which(grepl(condition.names[i], sampleNames(proteinGroups.CPTAC)))], breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(13,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("kNN imputation condition ", condition.names[i]), xlab = "log2(MaxLFQ intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(13,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
}

# Mixed imputation
for(i in 1:3){
p1 <- hist(exprs(proteinGroups.CPTAC.mixed)[,which(grepl(condition.names[i], sampleNames(proteinGroups.CPTAC.mixed)))], breaks=25, plot = FALSE)
p2 <- hist(exprs(proteinGroups.CPTAC)[,which(grepl(condition.names[i], sampleNames(proteinGroups.CPTAC)))], breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(13,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("Mixed imputation condition ", condition.names[i]), xlab = "log2(MaxLFQ intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(13,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
}

# RegEM imputation
for(i in 1:3){
p1 <- hist(exprs(proteinGroups.CPTAC.RegEM.imp)[,which(grepl(condition.names[i], sampleNames(proteinGroups.CPTAC.RegEM.imp)))], breaks=25, plot = FALSE)
p2 <- hist(exprs(proteinGroups.CPTAC)[,which(grepl(condition.names[i], sampleNames(proteinGroups.CPTAC)))], breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(13,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("RegEM imputation condition ", condition.names[i]), xlab = "log2(MaxLFQ intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(13,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
}
# dev.off()
```

## 3. Histograms for the effects of all imputations at peptide-level, grouped per condition (Supplementary Fig. 5)

```{r eval=TRUE}

condition.names <- unique(sampleNames(peptides.CPTAC.KNN1) %>% substr(3, 3))

cex.main <- 1.2
cex.lab <- 1
cex <- 1
par(mfrow=c(1,1))

# grDevices::svg(paste0(wd,"/histograms_all_imputations_peptides.svg"), width=60, height=120)
# cex.main <- 10
# cex.lab <- 8
# cex <- 6
# par(mfrow=c(4,3))
# par(mar=c(5.1*cex/2,4.1*cex/2,4.1*cex/2,2.1*cex/2))
# par(mfrow=c(9,3))

# Perseus imputation
for(i in 1:3){
p1 <- hist(exprs(peptides.CPTAC.Perseus.imp)[,which(grepl(condition.names[i], sampleNames(peptides.CPTAC.Perseus.imp)))], breaks=25, plot = FALSE)
p2 <- hist(exprs(peptides.CPTAC2)[,which(grepl(condition.names[i], sampleNames(peptides.CPTAC2)))], breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(13,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("Perseus imputation condition ", condition.names[i]), xlab = "log2(peptide intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(13,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
}

# kNN imputation
for(i in 1:3){
p1 <- hist(exprs(peptides.CPTAC.KNN1)[,which(grepl(condition.names[i], sampleNames(peptides.CPTAC.KNN1)))], breaks=25, plot = FALSE)
p2 <- hist(exprs(peptides.CPTAC2)[,which(grepl(condition.names[i], sampleNames(peptides.CPTAC2)))], breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(13,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("kNN imputation condition ", condition.names[i]), xlab = "log2(peptide intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(13,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
}

# Mixed imputation
for(i in 1:3){
p1 <- hist(exprs(peptides.CPTAC.mixed)[,which(grepl(condition.names[i], sampleNames(peptides.CPTAC.mixed)))], breaks=25, plot = FALSE)
p2 <- hist(exprs(peptides.CPTAC2)[,which(grepl(condition.names[i], sampleNames(peptides.CPTAC2)))], breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(13,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("Mixed imputation condition ", condition.names[i]), xlab = "log2(peptide intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(13,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
}

# RegEM imputation
for(i in 1:3){
p1 <- hist(exprs(peptides.CPTAC.RegEM.imp)[,which(grepl(condition.names[i], sampleNames(peptides.CPTAC.RegEM.imp)))], breaks=25, plot = FALSE)
p2 <- hist(exprs(peptides.CPTAC2)[,which(grepl(condition.names[i], sampleNames(peptides.CPTAC2)))], breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(13,28), border=rgb(255,165,0, maxColorValue = 255), las = 1, main = paste0("RegEM imputation condition ", condition.names[i]), xlab = "log2(peptide intensity)", cex.lab=cex.lab, cex.axis=cex, cex.main=cex.main, cex.sub=cex, lwd=cex)  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(13,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
}

# dev.off()
```

## Prepare for the FDR-FTP plots

```{r}
# Due to different preprocessing methods, some proteins cannot be detected with some methods.

# By default we use the full set, i.e. all protein identifiers over all methods.
# Note that all methods used either the preprocessing of MSqRob, of Perseus o of MSstats.
all.proteins <- unique(unlist(lapply(list(res.CPTAC.Hurdle, CPTAC.Perseus.full, res.CPTAC.MSstats.noimp.full), "[", "protein")))
full.Base <- tibble(protein = rep(all.proteins, 3), 
                    UPS = grepl("ups", protein),
                    contrast = rep(c("condition6B-condition6A",
                  "condition6C-condition6A",
                  "condition6C-condition6B"), each = length(all.proteins)))

# We will also make some figure with the common set, i.e. only the proteins that have been seen in the preprocessing of al methods.
common.proteins <- unique(unlist(lapply(list(res.CPTAC.Hurdle, CPTAC.Perseus.full, res.CPTAC.MSstats.noimp.full), "[", "protein")))


indices <- which(table(c(unique(as.character(res.CPTAC.Hurdle$protein)), unique(as.character(CPTAC.Perseus.full$protein)), unique(as.character(res.CPTAC.MSstats.noimp.full$protein)))) == 3)

common.proteins <- names(table(c(unique(as.character(res.CPTAC.Hurdle$protein)), unique(as.character(CPTAC.Perseus.full$protein)), unique(as.character(res.CPTAC.MSstats.noimp.full$protein)))))[indices]

common.Base <- tibble(protein = rep(common.proteins, 3), 
                    UPS = grepl("ups", protein),
                    contrast = rep(c("condition6B-condition6A",
                  "condition6C-condition6A",
                  "condition6C-condition6B"), each = length(common.proteins)))
```

## 4. Make the FDR-FTP plots for the methods that do not use Benjamini-Hochberg FDR correction as their default (i.e. Perseus and DEP) (Supplementary Fig. 5...6)

```{r}

# FDR-FTP plot for methods that do not use Benjamini-Hochberg FDR by default

res.list <- list(CPTAC.Perseus.imp.full,
                 CPTAC.Perseus.full,
                 CPTAC.Perseus.imp.full,
                 CPTAC.Perseus.full,
                 res.CPTAC.DEP,
                 res.CPTAC.DEP.noimp,
                 res.CPTAC.DEP,
                 res.CPTAC.DEP.noimp
)

res.list <- lapply(res.list, function(x){
    left_join(full.Base, x)
}
  )

contrast.vec <- c("condition6B-condition6A",
                  "condition6C-condition6A",
                  "condition6C-condition6B")

names(contrast.vec) <- c("Condition 6B vs. 6A",
                  "Condition 6C vs. 6A",
                  "Condition 6C vs. 6B")

colors <- c("green", "blue", "#42B7BA", "#1B2944", "darkgoldenrod", "#6B3D12", "peru", "yellow3")

sort.list <- list(c("q.BH", "p.value", "Test.statistic", "Difference"), # Benjamini-Hochberg FDR
                  c("q.BH", "p.value", "Test.statistic", "Difference"), # Benjamini-Hochberg FDR
                  c("q.value", "p.value", "Test.statistic", "Difference"), # permutation-based FDR
                  c("q.value", "p.value", "Test.statistic", "Difference"),  # permutation-based FDR
                  c("q.BH", "pvalue", "logFC"), # Benjamini-Hochberg FDR
                  c("q.BH", "pvalue", "logFC"), # Benjamini-Hochberg FDR
                  c("qvalue", "pvalue", "logFC"), # local fdr
                  c("qvalue", "pvalue", "logFC")) # local fdr

PlotFDPTPR(res.list, contrast.vec, colors, sort.list, TPcol = c("UPS"), plotSVG = FALSE) # TRUE
```

## 4. Make the FDR-FTP plots for the imputation-based methods (Supplementary Fig. 6)

```{r}

# FDR-FTP plot for all imputation-based methods

res.list <- list(res.CPTAC.full,
                 res.CPTAC.KNN1.full,
                 res.CPTAC.PI.full,
                 res.CPTAC.RegEM.full,
                 res.CPTAC.mixed.full
)

res.list <- lapply(res.list, function(x){
    left_join(full.Base, x)
}
  )

contrast.vec <- c("condition6B-condition6A",
                  "condition6C-condition6A",
                  "condition6C-condition6B")

names(contrast.vec) <- c("Condition 6B vs. 6A",
                  "Condition 6C vs. 6A",
                  "Condition 6C vs. 6B")

colors <- c("#FF681E", "#88FF9C", "blue", "black", "brown")

sort.list <- list(c("qvalue", "pvalue", "t", "logFC"),
                  c("qvalue", "pvalue", "t", "logFC"),
                  c("qvalue", "pvalue", "t", "logFC"),
                  c("qvalue", "pvalue", "t", "logFC"),
                  c("qvalue", "pvalue", "t", "logFC"))

PlotFDPTPR(res.list, contrast.vec, colors, sort.list, TPcol = c("UPS"), plotSVG = FALSE) # TRUE
```

## 5. Correlation between intensities and counts (Supplementary Fig. 7)

```{r}
# Make the plots

# BA, CB, CA
upratio <- c(1.565597, 1.571906, 3.137504)

# 1. Top: condition 6B vs. 6A

TP.BvsA <- res.CPTAC.Hurdle %>% filter(contrast == "condition6B-condition6A" & UPS)

FP.BvsA <- res.CPTAC.Hurdle %>% filter(contrast == "condition6B-condition6A" & !UPS)

scatterHist(x = ((FP.BvsA$logOR) %>% unlist),
            y = ((FP.BvsA$logFC) %>% unlist),
            x2 = ((TP.BvsA$logOR) %>% unlist),
            y2 = ((TP.BvsA$logFC) %>% unlist), 
            main = "Condition B vs. A", plotSVG = FALSE, filename = paste0(wd,"/scatterhistBA.svg")) #, plotSVG = TRUE, filename = paste0(wd,"/scatterhistBA.svg")

sum(is.na(((FP.BvsA$logFC) %>% unlist)))
# 159 yeast proteins without a log2FC estimate
sum(is.na(((TP.BvsA$logFC) %>% unlist)))
# 2 UPS1 proteins without a log2FC estimate

# 2. Middle: condition 6C vs. 6A

TP.CvsA <- res.CPTAC.Hurdle %>% filter(contrast == "condition6C-condition6A" & UPS)

FP.CvsA <- res.CPTAC.Hurdle %>% filter(contrast == "condition6C-condition6A" & !UPS)

scatterHist(x = ((FP.CvsA$logOR) %>% unlist),
            y = ((FP.CvsA$logFC) %>% unlist),
            x2 = ((TP.CvsA$logOR) %>% unlist),
            y2 = ((TP.CvsA$logFC) %>% unlist), 
            main = "Condition C vs. A", plotSVG = FALSE, filename = paste0(wd,"/scatterhistCA.svg")) #, plotSVG = TRUE, filename = paste0(wd,"/scatterhistCA.svg")

sum(is.na(((FP.CvsA$logFC) %>% unlist)))
# 158 yeast proteins without a log2FC estimate
sum(is.na(((TP.CvsA$logFC) %>% unlist)))
# 2 UPS1 proteins without a log2FC estimate

# 3. Bottom: condition 6C vs. 6B

TP.CvsB <- res.CPTAC.Hurdle %>% filter(contrast == "condition6C-condition6B" & UPS)

FP.CvsB <- res.CPTAC.Hurdle %>% filter(contrast == "condition6C-condition6B" & !UPS)

scatterHist(x = ((FP.CvsB$logOR) %>% unlist),
            y = ((FP.CvsB$logFC) %>% unlist),
            x2 = ((TP.CvsB$logOR) %>% unlist),
            y2 = ((TP.CvsB$logFC) %>% unlist), 
            main = "Condition C vs. B", plotSVG = FALSE, filename = paste0(wd,"/scatterhistCB.svg")) #, plotSVG = TRUE, filename = paste0(wd,"/scatterhistCB.svg")

sum(is.na(((FP.CvsB$logFC) %>% unlist)))
# 159 yeast proteins without a log2FC estimate
sum(is.na(((TP.CvsB$logFC) %>% unlist)))
# 2 UPS1 proteins without a log2FC estimate
```

## 6. Make the FDR-FTP plots for the count-based methods (Supplementary Fig. 8)

```{r}

# FDR-FTP plots for the count-based methods

res.list <- list(res.CPTAC.sc.co.full,
                 res.CPTAC.edgeR.sc.full,
                 res.CPTAC.beta.binom.sc.full,
                 res.CPTAC.co.full,
                 res.CPTAC.edgeR.full,
                 res.CPTAC.beta.binom.full,
                 res.CPTAC.full
)

res.list <- lapply(res.list, function(x){
  left_join(full.Base, x)
}
)

contrast.vec <- c("condition6B-condition6A",
                  "condition6C-condition6A",
                  "condition6C-condition6B")

names(contrast.vec) <- c("Condition 6B vs. 6A",
                         "Condition 6C vs. 6A",
                         "Condition 6C vs. 6B")

colors <- c("magenta", "lightgreen", "gold1", "#E15E9E", "darkgreen","yellow3", "#FF681E")

sort.list <- list(c("qvalue", "pvalue", "t", "logOR"),
                  c("FDR", "PValue", "F", "logFC"),
                  c("qvalue", "p.value"), 
                  c("qvalue", "pvalue", "t", "logOR"),
                  c("FDR", "PValue", "F", "logFC"),
                  c("qvalue", "p.value"),
                  c("qvalue", "pvalue", "t", "logFC"))

PlotFDPTPR(res.list, contrast.vec, colors, sort.list, TPcol = c("UPS"), plotSVG = FALSE) # TRUE
```

## 7. On the independence of the z-statistics under the null hypothesis  (Supplementary Fig. 9)

```{r}
# Create a mock experimental design where there is no differential abundance in reality
# There is however a huge "batch effect" of "condlab", the combined effects of spike-in condition and lab
pData <- pData(peptides.CPTAC2)
pData$condlab <- paste0(pData$condition,pData$lab)
pData$treat <- rep(c("one","two","three"),9)
pData$treat <- factor(pData$treat, levels = c("one", "two", "three"))

peptides.CPTAC2.noeff <- MSnSet(Biobase::exprs(peptides.CPTAC2), fData = AnnotatedDataFrame(fData(peptides.CPTAC2)), pData = AnnotatedDataFrame(pData))

### Recreate peptides.CPTAC3 ###

# We require by default at least 2 identifications of a peptide sequence, as with 1 identification, the model will be perfectly confounded
sel <- rowSums(!is.na(Biobase::exprs(peptides.CPTAC2))) >= 2
peptides.CPTAC3 <- peptides.CPTAC2[sel]

# Again remove proteins that are only identified by one peptide
sel <- fData(peptides.CPTAC3) %>% group_by(protein) %>% mutate(flag = n() > 1) %>% pull(flag)
peptides.CPTAC3 <- peptides.CPTAC3[sel]

# Drop levels
peptides.CPTAC3 <- MSnbase::MSnSet(exprs = Biobase::exprs(peptides.CPTAC3), fData = droplevels(Biobase::fData(peptides.CPTAC3)), pData = Biobase::pData(peptides.CPTAC3))

dim(peptides.CPTAC2)
# 9377   27
dim(peptides.CPTAC3)
# 9158   27
length(unique(fData(peptides.CPTAC2)$protein))
# 1381
length(unique(fData(peptides.CPTAC3)$protein))
# 1347

rm(sel)
gc()
######

peptides.CPTAC3.noeff <- MSnSet(Biobase::exprs(peptides.CPTAC3), fData = AnnotatedDataFrame(fData(peptides.CPTAC3)), pData = AnnotatedDataFrame(pData))

# Create contrast matrix for the mock analysis
form <- expression ~ (1|treat) + (1|condlab) + (1|sample) + (1|sequence)
contrasts <- contrast_helper(form, peptides.CPTAC3.noeff, treat)
contrasts <- contrasts[c(1,3,2),c(2,1,3)]
colnames(contrasts)[3] <- "treatthree-treattwo"
contrasts[c(2,3),3] <- c(-1,1)
contrasts

# Fit the models, this might take a few minutes
# res.CPTAC.noeff <- do_mm(formula = form, msnset = peptides.CPTAC3.noeff, group_var = protein, contrasts = contrasts, type_df = "conservative", max_iter = 20)
# Or load the data
load(paste0(wd, "/save_files_mock_CPTAC/res_CPTAC_noeff.RData"))

annotation.CPTAC <- tibble(protein = fData(peptides.CPTAC2)$protein, UPS = grepl("ups",fData(peptides.CPTAC2)$protein), gene.name = fData(peptides.CPTAC2)$gene.name, protein.name = fData(peptides.CPTAC2)$protein.name) %>% distinct
res.CPTAC.Base <- rbind(annotation.CPTAC, annotation.CPTAC, annotation.CPTAC)
res.CPTAC.Base$contrast <- c(rep("treattwo-treatone",nrow(annotation.CPTAC)), rep("treatthree-treatone",nrow(annotation.CPTAC)), rep("treatthree-treattwo",nrow(annotation.CPTAC)))

res.CPTAC.noeff.full <- res.CPTAC.Base %>% left_join(res.CPTAC.noeff$result)
res.CPTAC.noeff.full <- res.CPTAC.noeff.full %>% arrange(pvalue)

prot.CPTAC.noeff.co <- do_count_groups(peptides.CPTAC2.noeff, group_var = protein, keep_fData_cols = c("gene.name","protein.name"))

# Remove estimates based on only one sample in one or more conditions from the data

not_one <- which((Biobase::exprs(prot.CPTAC.noeff.co)[,seq(1,25, by=3)] > 0) %>% rowSums < 2)
not_two <- which((Biobase::exprs(prot.CPTAC.noeff.co)[,seq(2,26, by=3)] > 0) %>% rowSums < 2)
not_three <- which((Biobase::exprs(prot.CPTAC.noeff.co)[,seq(3,27, by=3)] > 0) %>% rowSums < 2)

### Important: re-adjust the false discovery rate! ###

res.CPTAC.noeff.full[(res.CPTAC.noeff.full$protein %in% names(not_one)) & (res.CPTAC.noeff.full$contrast %in% c("treattwo-treatone", "treatthree-treatone")),][,c("logFC", "se", "t", "df", "pvalue", "qvalue")] <- NA
res.CPTAC.noeff.full[(res.CPTAC.noeff.full$protein %in% names(not_two)) & (res.CPTAC.noeff.full$contrast %in% c("treattwo-treatone", "treatthree-treattwo")),][,c("logFC", "se", "t", "df", "pvalue", "qvalue")] <- NA
res.CPTAC.noeff.full[(res.CPTAC.noeff.full$protein %in% names(not_three)) & (res.CPTAC.noeff.full$contrast %in% c("treatthree-treatone", "treatthree-treattwo")),][,c("logFC", "se", "t", "df", "pvalue", "qvalue")] <- NA

res.CPTAC.noeff.full <- res.CPTAC.noeff.full %>% group_by(contrast) %>% 
  mutate(qvalue = p.adjust(pvalue, method = "BH"))

# Or load the MSqRob result object via:
# load(paste0(wd, "/save_files_mock_CPTAC/res_CPTAC_noeff_full.RData"))

### Fit the quasi-binomial model ###

form.co <- ~ treat + condlab
# contrasts <- contrast_helper(~ treat, peptides.CPTAC2.noeff, treat)

# res.CPTAC.noeff.co <- do_glm(formula = form.co, msnset = prot.CPTAC.noeff.co, group_var = protein, contrasts = contrasts, contFun = "contEst")
# OR: load the model:
load(paste0(wd, "/save_files_mock_CPTAC/res_CPTAC_noeff_co.RData"))

res.CPTAC.noeff.co.full <- res.CPTAC.Base %>% left_join(res.CPTAC.noeff.co$result)
res.CPTAC.noeff.co.full <- res.CPTAC.noeff.co.full %>% arrange(pvalue)

# cols <-  colorRampPalette(c("#FF3100", "#FCFF00", "#45FE4F",
#                             "#00FEFF", "#000099" 
# ))(256)

# col = cols[ceiling(rank(HEART_A_vsV$pchisq)/length(HEART_A_vsV$pchisq)*256)]

# Order results according to protein and contrast instead of p-value
res.intensities.sorted <- res.CPTAC.noeff.full %>% arrange(contrast, protein)
res.counts.sorted <- res.CPTAC.noeff.co.full %>% arrange(contrast, protein)

all(res.intensities.sorted$protein == res.counts.sorted$protein)
all(res.intensities.sorted$contrast == res.counts.sorted$contrast)

# library(svglite)
# svglite(file = "all_mock_correlations.svg", width = 14, height = 14)
# # svglite(file = "no_correlation_MSqRob_qbinom_21.svg", width = 7, height = 7)
# par(mar=c(5.1,6.1,4.1,2.1))
# par(mfrow = c(2,2))
plot(res.intensities.sorted[res.intensities.sorted$contrast == "treattwo-treatone",]$logFC, res.counts.sorted[res.counts.sorted$contrast == "treattwo-treatone",]$logOR, col = c("black","red")[as.numeric(res.intensities.sorted$UPS)+1], pch = 20, xlab = "MSqRob log2FC", ylab = "Quasi-binomial log2OR", las = 1, frame.plot = FALSE, main = "Nonsense treatment 2 vs. 1", cex.lab = 2, cex.axis = 2, cex.main = 2) #[1:1000]
abline(v = 0, col = "black")
abline(h = 0, col = "black")
# # par(mar=c(5.1,4.1,4.1,2.1))
# # dev.off()
# 
# # library(svglite)
# # svglite(file = "no_correlation_MSqRob_qbinom_31.svg", width = 7, height = 7)
# # par(mar=c(5.1,6.1,4.1,2.1))
plot(res.intensities.sorted[res.intensities.sorted$contrast == "treatthree-treatone",]$logFC, res.counts.sorted[res.counts.sorted$contrast == "treatthree-treatone",]$logOR, col = c("black","red")[as.numeric(res.intensities.sorted$UPS)+1], pch = 20, xlab = "MSqRob log2FC", ylab = "Quasi-binomial log2OR", las = 1, frame.plot = FALSE, main = "Nonsense treatment 3 vs. 1", cex.lab = 2, cex.axis = 2, cex.main = 2) #[1:1000]
abline(v = 0, col = "black")
abline(h = 0, col = "black")
# # par(mar=c(5.1,4.1,4.1,2.1))
# # dev.off()
# 
# # library(svglite)
# # svglite(file = "no_correlation_MSqRob_qbinom_32.svg", width = 7, height = 7)
# # par(mar=c(5.1,6.1,4.1,2.1))
plot(res.intensities.sorted[res.intensities.sorted$contrast == "treatthree-treattwo",]$logFC, res.counts.sorted[res.counts.sorted$contrast == "treatthree-treattwo",]$logOR, col = c("black","red")[as.numeric(res.intensities.sorted$UPS)+1], pch = 20, xlab = "MSqRob log2FC", ylab = "Quasi-binomial log2OR", las = 1, frame.plot = FALSE, main = "Nonsense treatment 3 vs. 2", cex.lab = 2, cex.axis = 2, cex.main = 2) #[1:1000]
abline(v = 0, col = "black")
abline(h = 0, col = "black")
# # par(mar=c(5.1,4.1,4.1,2.1))
# par(mfrow = c(1,1))
# dev.off()
```

## 10. For the common set, make the FDR-FTP plots for maxLFQ-based state-of-the-art methods (Supplementary Fig. 12)

```{r}
res.list <- list(CPTAC.Perseus.imp.full,
     CPTAC.Perseus.full,
     res.CPTAC.DEP,
     res.CPTAC.mixture,
     res.CPTAC.ProDA.full,
     res.CPTAC.DEP.noimp,
     res.CPTAC.co.full,
     res.CPTAC.full,
     res.CPTAC.Hurdle)

res.list <- lapply(res.list, function(x){
    left_join(common.Base, x)
}
  )

contrast.vec <- c("condition6B-condition6A",
                  "condition6C-condition6A",
                  "condition6C-condition6B")

names(contrast.vec) <- c("Condition 6B vs. 6A",
                  "Condition 6C vs. 6A",
                  "Condition 6C vs. 6B")

colors <- c("#42B7BA", "#1B2944", "peru", "#FFC30B", "#787878", "yellow3", "#E15E9E", "#FF681E", "#5A2A82")

sort.list <- list(c("q.value", "p.value", "Test.statistic", "Difference"),
     c("q.value", "p.value", "Test.statistic", "Difference"),
     c("qvalue", "pvalue", "logFC"),
     c("qvalue", "pvalue"),
     c("qval", "pval", "t", "logFC"),
     c("qvalue", "pvalue", "logFC"),
     c("qvalue", "pvalue", "t", "logOR"),
     c("qvalue", "pvalue", "t", "logFC"),
     c("qchisq", "pchisq", "chisq", NA))

PlotFDPTPR(res.list, contrast.vec, colors, sort.list, TPcol = c("UPS"), plotSVG = FALSE) # TRUE
```

## 3. For the common set, make the FDR-FTP plots for peptide-based state-of-the-art methods methods

```{r eval = TRUE}
res.list <- list(res.CPTAC.MSstats.full[!is.infinite(res.CPTAC.MSstats.full$log2FC),],
                 res.CPTAC.MSstats.noimp.full[!is.infinite(res.CPTAC.MSstats.noimp.full$log2FC),],
                 res.CPTAC.ProPCA.full.QN.int, 
                 res.CPTAC.RegEM.full,
                 res.CPTAC.co.full,
                 res.CPTAC.full,
                 res.CPTAC.Hurdle)

res.list <- lapply(res.list, function(x){
    left_join(common.Base, x)
}
  )

contrast.vec <- c("condition6B-condition6A",
                  "condition6C-condition6A",
                  "condition6C-condition6B")

names(contrast.vec) <- c("Condition 6B vs. 6A",
                  "Condition 6C vs. 6A",
                  "Condition 6C vs. 6B")

colors <- c("#808000", "green", "red2", "black", "#E15E9E", "#FF681E", "#5A2A82") # darkblue: "#1B2944"  gray: "#50FF00"

sort.list <- list(c("adj.pvalue", "pvalue", "Tvalue", "log2FC"),
                  c("adj.pvalue", "pvalue", "Tvalue", "log2FC"),
                  c("adj.P.Val", "P.Value", "t", "logFC"), 
                  c("qvalue", "pvalue", "t", "logFC"),
                  c("qvalue", "pvalue", "t", "logOR"),
                  c("qvalue", "pvalue", "t", "logFC"),
                  c("qchisq", "pchisq", "chisq", NA))

PlotFDPTPR(res.list, contrast.vec, colors, sort.list, TPcol = c("UPS"), plotSVG = FALSE) # TRUE
```

## 11. Plot example proteins for each of the overlapping regions in the Venn diagram (Supplementary Fig. 13 - 17)

Aditionally, we here provide the plots for the underlying (preprocessed) peptide-level data for the 1500 most significant proteins for each method in the HEART dataset, grouped per method.

```{r}

### First 1500 most significant genes ###

# A. Select top 1500 DA gene identifiers for the Perseus analysis of ventricles vs. atria
genes.Perseus.AvsV.1500 <- (Perseus.AvsV.ov %>% pull(`Gene names`) %>% unique)[1:1500]

# B. Select top 1500 DA gene identifiers for the Hurdle analysis of atria vs. ventricles
genes.Hurdle.AvsV.1500 <- hurdle.AvsV.ov[1:1500,]$gene.name

# C. Select top 1500 DA gene identifiers for the MSqRob analysis of atria vs. ventricles
genes.MSqRob.AvsV.1500 <- MSqRob.AvsV.ov[1:1500,]$gene.name


### Overlap all (Supplementary Fig. 8) ###

plot_proteins(gene.names = "MYL7", title = "MYL7", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "PAM", title = "PAM", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "MYBPHL", title = "MYBPHL", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "MAP1B", title = "MAP1B", msnset = peptides.HEART2, pdf = FALSE)

# plot_proteinsSVG(gene.names = c("MYL7", "PAM", "MYBPHL", "MAP1B"), title = c("overlap_all.svg"), msnset = peptides.HEART2)
# Order p-values by evidence from all three methods
genes <- genes.MSqRob.AvsV.1500[(genes.MSqRob.AvsV.1500 %in% genes.Perseus.AvsV.1500) & (genes.MSqRob.AvsV.1500 %in% genes.Hurdle.AvsV.1500)] %>% as.character

Mdata <- MSqRob.AvsV.ov[MSqRob.AvsV.ov$gene.name %in% genes,]

Pdata <- Perseus.AvsV.ov[Perseus.AvsV.ov$`Gene names` %in% genes,]

Hdata <- hurdle.AvsV.ov[hurdle.AvsV.ov$gene.name %in% genes,]

MSqRob.pvals <- Mdata$pvalue
names(MSqRob.pvals) <- Mdata$gene.name
MSqRob.pvals <- MSqRob.pvals[genes]
MSqRob.pvals[is.na(MSqRob.pvals)] <- 1

Perseus.pvals <- 10^(-Pdata$`minusLOG(P-value)`)
names(Perseus.pvals) <- Pdata$`Gene names`
Perseus.pvals <- Perseus.pvals[genes]
Perseus.pvals[is.na(Perseus.pvals)] <- 1

hurdle.pvals <- Hdata$pchisq
names(hurdle.pvals) <- Hdata$gene.name
hurdle.pvals <- hurdle.pvals[genes]
hurdle.pvals[is.na(hurdle.pvals)] <- 1

genes.ordered <- qnorm(MSqRob.pvals/2)^2+qnorm(Perseus.pvals/2)^2+qnorm(hurdle.pvals/2)^2
genes.ordered <- sort(genes.ordered, decreasing = TRUE)
# plot_proteins(gene.names = names(genes.ordered), title = "overlap_all.pdf", msnset = peptides.HEART2, pdf = TRUE)


### Overlap Hurdle and Perseus (Supplementary Fig. 9) ###

plot_proteins(gene.names = "CA13", title = "CA13", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "NTN1", title = "NTN1", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "PRKG2", title = "PRKG2", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "SBF2", title = "SBF2", msnset = peptides.HEART2, pdf = FALSE)

# plot_proteinsSVG(gene.names = c("CA13", "NTN1", "PRKG2", "SBF2"), title = c("overlap_hurdle_Perseus.svg"), msnset = peptides.HEART2)
# Order p-values by evidence from Hurdle and Perseus and against MSqRob
genes <- genes.Hurdle.AvsV.1500[(genes.Hurdle.AvsV.1500 %in% genes.Perseus.AvsV.1500) & !(genes.Hurdle.AvsV.1500 %in% genes.MSqRob.AvsV.1500)] %>% as.character

Mdata <- MSqRob.AvsV.ov[MSqRob.AvsV.ov$gene.name %in% genes,]

Pdata <- Perseus.AvsV.ov[Perseus.AvsV.ov$`Gene names` %in% genes,]

Hdata <- hurdle.AvsV.ov[hurdle.AvsV.ov$gene.name %in% genes,]

MSqRob.pvals <- Mdata$pvalue
names(MSqRob.pvals) <- Mdata$gene.name
MSqRob.pvals <- MSqRob.pvals[genes]
MSqRob.pvals[is.na(MSqRob.pvals)] <- 1

Perseus.pvals <- 10^(-Pdata$`minusLOG(P-value)`)
names(Perseus.pvals) <- Pdata$`Gene names`
Perseus.pvals <- Perseus.pvals[genes]
Perseus.pvals[is.na(Perseus.pvals)] <- 1

hurdle.pvals <- Hdata$pchisq
names(hurdle.pvals) <- Hdata$gene.name
hurdle.pvals <- hurdle.pvals[genes]
hurdle.pvals[is.na(hurdle.pvals)] <- 1

genes.ordered <- qnorm((1-MSqRob.pvals)/2)^2+qnorm(Perseus.pvals/2)^2+qnorm(hurdle.pvals/2)^2
genes.ordered <- sort(genes.ordered, decreasing = TRUE)
# plot_proteins(gene.names = names(genes.ordered), title = "overlap_hurdle_Perseus.pdf", msnset = peptides.HEART2, pdf = TRUE)


### Hurdle alone (Supplementary Fig. 10) ###

plot_proteins(gene.names = "HTT", title = "HTT", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "ACACA", title = "ACACA", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "ABCA6", title = "ABCA6", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "FTCD", title = "FTCD", msnset = peptides.HEART2, pdf = FALSE)

# plot_proteinsSVG(gene.names = c("HTT", "ACACA", "ABCA6", "FTCD"), title = c("only_hurdle.svg"), msnset = peptides.HEART2)
# Order p-values by evidence from Hurdle and against MSqRob and Perseus
genes <- genes.Hurdle.AvsV.1500[!(genes.Hurdle.AvsV.1500 %in% genes.MSqRob.AvsV.1500) & !(genes.Hurdle.AvsV.1500 %in% genes.Perseus.AvsV.1500)] %>% as.character

Mdata <- MSqRob.AvsV.ov[MSqRob.AvsV.ov$gene.name %in% genes,]

Pdata <- Perseus.AvsV.ov[Perseus.AvsV.ov$`Gene names` %in% genes,]

Hdata <- hurdle.AvsV.ov[hurdle.AvsV.ov$gene.name %in% genes,]

MSqRob.pvals <- Mdata$pvalue
names(MSqRob.pvals) <- Mdata$gene.name
MSqRob.pvals <- MSqRob.pvals[genes]
MSqRob.pvals[is.na(MSqRob.pvals)] <- 1

Perseus.pvals <- 10^(-Pdata$`minusLOG(P-value)`)
names(Perseus.pvals) <- Pdata$`Gene names`
Perseus.pvals <- Perseus.pvals[genes]
Perseus.pvals[is.na(Perseus.pvals)] <- 1

hurdle.pvals <- Hdata$pchisq
names(hurdle.pvals) <- Hdata$gene.name
hurdle.pvals <- hurdle.pvals[genes]
hurdle.pvals[is.na(hurdle.pvals)] <- 1

genes.ordered <- qnorm((1-MSqRob.pvals)/2)^2+qnorm((1-Perseus.pvals)/2)^2+qnorm(hurdle.pvals/2)^2
genes.ordered <- sort(genes.ordered, decreasing = TRUE)
# plot_proteins(gene.names = names(genes.ordered), title = "only_hurdle.pdf", msnset = peptides.HEART2, pdf = TRUE)


### Perseus alone (Supplementary Fig. 11) ###

plot_proteins(gene.names = "GJA5", title = "GJA5", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "ASNSD1", title = "ASNSD1", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "MTCL1", title = "MTCL1", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "BLOC1S5;BLOC1S5-TXNDC5", title = "BLOC1S5;BLOC1S5-TXNDC5", msnset = peptides.HEART2, pdf = FALSE)

# plot_proteinsSVG(gene.names = c("GJA5", "ASNSD1", "MTCL1", "BLOC1S5;BLOC1S5-TXNDC5"), title = c("only_Perseus.svg"), msnset = peptides.HEART2)
# Order p-values by evidence from Perseus and against Hurdle and MSqRob
genes <- genes.Perseus.AvsV.1500[!(genes.Perseus.AvsV.1500 %in% genes.MSqRob.AvsV.1500) & !(genes.Perseus.AvsV.1500 %in% genes.Hurdle.AvsV.1500)] %>% as.character

Mdata <- MSqRob.AvsV.ov[MSqRob.AvsV.ov$gene.name %in% genes,]

Pdata <- Perseus.AvsV.ov[Perseus.AvsV.ov$`Gene names` %in% genes,]

Hdata <- hurdle.AvsV.ov[hurdle.AvsV.ov$gene.name %in% genes,]

MSqRob.pvals <- Mdata$pvalue
names(MSqRob.pvals) <- Mdata$gene.name
MSqRob.pvals <- MSqRob.pvals[genes]
MSqRob.pvals[is.na(MSqRob.pvals)] <- 1

Perseus.pvals <- 10^(-Pdata$`minusLOG(P-value)`)
names(Perseus.pvals) <- Pdata$`Gene names`
Perseus.pvals <- Perseus.pvals[genes]
Perseus.pvals[is.na(Perseus.pvals)] <- 1

hurdle.pvals <- Hdata$pchisq
names(hurdle.pvals) <- Hdata$gene.name
hurdle.pvals <- hurdle.pvals[genes]
hurdle.pvals[is.na(hurdle.pvals)] <- 1

genes.ordered <- qnorm((1-MSqRob.pvals)/2)^2+qnorm(Perseus.pvals/2)^2+qnorm((1-hurdle.pvals)/2)^2
genes.ordered <- sort(genes.ordered, decreasing = TRUE)
# plot_proteins(gene.names = names(genes.ordered), title = "only_Perseus.pdf", msnset = peptides.HEART2, pdf = TRUE)


### Overlap MSqRob and Perseus (Supplementary Fig. 12) ###

plot_proteins(gene.names = "AK4", title = "AK4", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "NDUFV1", title = "NDUFV1", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "DECR1", title = "DECR1", msnset = peptides.HEART2, pdf = FALSE)
plot_proteins(gene.names = "SDHA", title = "SDHA", msnset = peptides.HEART2, pdf = FALSE)

# plot_proteinsSVG(gene.names = c("AK4", "NDUFV1", "DECR1", "SDHA"), title = c("overlap_MSqRob_Perseus.svg"), msnset = peptides.HEART2)
# Order p-values by evidence from MSqRob and Perseus and against Hurdle
genes <- genes.MSqRob.AvsV.1500[(genes.MSqRob.AvsV.1500 %in% genes.Perseus.AvsV.1500) & !(genes.MSqRob.AvsV.1500 %in% genes.Hurdle.AvsV.1500)] %>% as.character

Mdata <- MSqRob.AvsV.ov[MSqRob.AvsV.ov$gene.name %in% genes,]

Pdata <- Perseus.AvsV.ov[Perseus.AvsV.ov$`Gene names` %in% genes,]

Hdata <- hurdle.AvsV.ov[hurdle.AvsV.ov$gene.name %in% genes,]

MSqRob.pvals <- Mdata$pvalue
names(MSqRob.pvals) <- Mdata$gene.name
MSqRob.pvals <- MSqRob.pvals[genes]
MSqRob.pvals[is.na(MSqRob.pvals)] <- 1

Perseus.pvals <- 10^(-Pdata$`minusLOG(P-value)`)
names(Perseus.pvals) <- Pdata$`Gene names`
Perseus.pvals <- Perseus.pvals[genes]
Perseus.pvals[is.na(Perseus.pvals)] <- 1

hurdle.pvals <- Hdata$pchisq
names(hurdle.pvals) <- Hdata$gene.name
hurdle.pvals <- hurdle.pvals[genes]
hurdle.pvals[is.na(hurdle.pvals)] <- 1

genes.ordered <- qnorm(MSqRob.pvals/2)^2+qnorm(Perseus.pvals/2)^2+qnorm((1-hurdle.pvals)/2)^2
genes.ordered <- sort(genes.ordered, decreasing = TRUE)
# plot_proteins(gene.names = names(genes.ordered), title = "overlap_MSqRob_Perseus.pdf", msnset = peptides.HEART2, pdf = TRUE)


### MSqRob alone ###

# Order p-values by evidence from MSqRob and against Hurdle and Perseus
genes <- genes.MSqRob.AvsV.1500[!(genes.MSqRob.AvsV.1500 %in% genes.Perseus.AvsV.1500) & !(genes.MSqRob.AvsV.1500 %in% genes.Hurdle.AvsV.1500)] %>% as.character

Mdata <- MSqRob.AvsV.ov[MSqRob.AvsV.ov$gene.name %in% genes,]

Pdata <- Perseus.AvsV.ov[Perseus.AvsV.ov$`Gene names` %in% genes,]

Hdata <- hurdle.AvsV.ov[hurdle.AvsV.ov$gene.name %in% genes,]

MSqRob.pvals <- Mdata$pvalue
names(MSqRob.pvals) <- Mdata$gene.name
MSqRob.pvals <- MSqRob.pvals[genes]
MSqRob.pvals[is.na(MSqRob.pvals)] <- 1

Perseus.pvals <- 10^(-Pdata$`minusLOG(P-value)`)
names(Perseus.pvals) <- Pdata$`Gene names`
Perseus.pvals <- Perseus.pvals[genes]
Perseus.pvals[is.na(Perseus.pvals)] <- 1

hurdle.pvals <- Hdata$pchisq
names(hurdle.pvals) <- Hdata$gene.name
hurdle.pvals <- hurdle.pvals[genes]
hurdle.pvals[is.na(hurdle.pvals)] <- 1

genes.ordered <- qnorm(MSqRob.pvals/2)^2+qnorm((1-Perseus.pvals)/2)^2+qnorm((1-hurdle.pvals)/2)^2
genes.ordered <- sort(genes.ordered, decreasing = TRUE)
# plot_proteins(gene.names = names(genes.ordered), title = "only_MSqRob.pdf", msnset = peptides.HEART2, pdf = TRUE)


### Overlap MSqRob and Hurdle ###

# Order p-values by evidence from MSqRob and Hurdle and against Perseus
genes <- genes.MSqRob.AvsV.1500[(genes.MSqRob.AvsV.1500 %in% genes.Hurdle.AvsV.1500) & !(genes.MSqRob.AvsV.1500 %in% genes.Perseus.AvsV.1500)] %>% as.character

Mdata <- MSqRob.AvsV.ov[MSqRob.AvsV.ov$gene.name %in% genes,]

Pdata <- Perseus.AvsV.ov[Perseus.AvsV.ov$`Gene names` %in% genes,]

Hdata <- hurdle.AvsV.ov[hurdle.AvsV.ov$gene.name %in% genes,]

MSqRob.pvals <- Mdata$pvalue
names(MSqRob.pvals) <- Mdata$gene.name
MSqRob.pvals <- MSqRob.pvals[genes]
MSqRob.pvals[is.na(MSqRob.pvals)] <- 1

Perseus.pvals <- 10^(-Pdata$`minusLOG(P-value)`)
names(Perseus.pvals) <- Pdata$`Gene names`
Perseus.pvals <- Perseus.pvals[genes]
Perseus.pvals[is.na(Perseus.pvals)] <- 1

hurdle.pvals <- Hdata$pchisq
names(hurdle.pvals) <- Hdata$gene.name
hurdle.pvals <- hurdle.pvals[genes]
hurdle.pvals[is.na(hurdle.pvals)] <- 1

genes.ordered <- qnorm(MSqRob.pvals/2)^2+qnorm((1-Perseus.pvals)/2)^2+qnorm(hurdle.pvals/2)^2
genes.ordered <- sort(genes.ordered, decreasing = TRUE)
# plot_proteins(gene.names = names(genes.ordered), title = "overlap_MSqRob_Hurdle.pdf", msnset = peptides.HEART2, pdf = TRUE)

```

## 12. Generate the data for the other Venn diagrams for the HEART dataset (Supplementary Fig. 18)

```{r}
### Venn diagrams atrial vs. ventricular region ###

### Checks: ###

all(Perseus.AvsV.ov$`Gene names` %in% hurdle.AvsV.ov$gene.name)
all(Perseus.AvsV.ov$`Gene names` %in% MSqRob.AvsV.ov$gene.name)

all(hurdle.AvsV.ov$gene.name %in% Perseus.AvsV.ov$`Gene names`)
all(MSqRob.AvsV.ov$gene.name %in% Perseus.AvsV.ov$`Gene names`)

dim(MSqRob.AvsV.ov)
dim(hurdle.AvsV.ov)
length(unique(Perseus.AvsV.ov$`Gene names`))

################

### Top: the first 500 significant genes ###

# A. Select top 500 DA gene identifiers for the Perseus analysis of ventricles vs. atria
genes.Perseus.AvsV.500 <- (Perseus.AvsV.ov %>% pull(`Gene names`) %>% unique)[1:500]

# B. Select top 500 DA gene identifiers for the Hurdle analysis of atria vs. ventricles
genes.Hurdle.AvsV.500 <- hurdle.AvsV.ov[1:500,]$gene.name

# C. Select top 500 DA gene identifiers for the MSqRob analysis of atria vs. ventricles
genes.MSqRob.AvsV.500 <- MSqRob.AvsV.ov[1:500,]$gene.name

# MSqRob alone

sum(!(genes.MSqRob.AvsV.500 %in% genes.Hurdle.AvsV.500) & !(genes.MSqRob.AvsV.500 %in% genes.Perseus.AvsV.500))
# 99

# Hurdle alone

sum(!(genes.Hurdle.AvsV.500 %in% genes.MSqRob.AvsV.500) & !(genes.Hurdle.AvsV.500 %in% genes.Perseus.AvsV.500))
# 85

# Perseus alone

sum(!(genes.Perseus.AvsV.500 %in% genes.MSqRob.AvsV.500) & !(genes.Perseus.AvsV.500 %in% genes.Hurdle.AvsV.500))
# 218

# Overlap MSqRob and Hurdle

sum((genes.MSqRob.AvsV.500 %in% genes.Hurdle.AvsV.500) & !(genes.MSqRob.AvsV.500 %in% genes.Perseus.AvsV.500))
# 158

# Overlap MSqRob and Perseus

sum((genes.MSqRob.AvsV.500 %in% genes.Perseus.AvsV.500) & !(genes.MSqRob.AvsV.500 %in% genes.Hurdle.AvsV.500))
# 25

# Overlap Hurdle and Perseus

sum((genes.Hurdle.AvsV.500 %in% genes.Perseus.AvsV.500) & !(genes.Hurdle.AvsV.500 %in% genes.MSqRob.AvsV.500))
# 39

# Overlap everything

sum((genes.MSqRob.AvsV.500 %in% genes.Perseus.AvsV.500) & (genes.MSqRob.AvsV.500 %in% genes.Hurdle.AvsV.500))
# 218

# Control:
length(genes.MSqRob.AvsV.500)
99+218+158+25

length(genes.Hurdle.AvsV.500)
85+158+218+39

length(genes.Perseus.AvsV.500)
218+39+25+218

### Bottom: the first 1000 significant genes ###

# A. Select top 1000 DA gene identifiers for the Perseus analysis of ventricles vs. atria
genes.Perseus.AvsV.1000 <- (Perseus.AvsV.ov %>% pull(`Gene names`) %>% unique)[1:1000]

# B. Select top 1000 DA gene identifiers for the Hurdle analysis of atria vs. ventricles
genes.Hurdle.AvsV.1000 <- hurdle.AvsV.ov[1:1000,]$gene.name

# C. Select top 1000 DA gene identifiers for the MSqRob analysis of atria vs. ventricles
genes.MSqRob.AvsV.1000 <- MSqRob.AvsV.ov[1:1000,]$gene.name

# MSqRob alone

sum(!(genes.MSqRob.AvsV.1000 %in% genes.Hurdle.AvsV.1000) & !(genes.MSqRob.AvsV.1000 %in% genes.Perseus.AvsV.1000))
# 174

# Hurdle alone

sum(!(genes.Hurdle.AvsV.1000 %in% genes.MSqRob.AvsV.1000) & !(genes.Hurdle.AvsV.1000 %in% genes.Perseus.AvsV.1000))
# 134

# Perseus alone

sum(!(genes.Perseus.AvsV.1000 %in% genes.MSqRob.AvsV.1000) & !(genes.Perseus.AvsV.1000 %in% genes.Hurdle.AvsV.1000))
# 439

# Overlap MSqRob and Hurdle

sum((genes.MSqRob.AvsV.1000 %in% genes.Hurdle.AvsV.1000) & !(genes.MSqRob.AvsV.1000 %in% genes.Perseus.AvsV.1000))
# 363

# Overlap MSqRob and Perseus

sum((genes.MSqRob.AvsV.1000 %in% genes.Perseus.AvsV.1000) & !(genes.MSqRob.AvsV.1000 %in% genes.Hurdle.AvsV.1000))
# 58

# Overlap Hurdle and Perseus

sum((genes.Hurdle.AvsV.1000 %in% genes.Perseus.AvsV.1000) & !(genes.Hurdle.AvsV.1000 %in% genes.MSqRob.AvsV.1000))
# 98

# Overlap everything

sum((genes.MSqRob.AvsV.1000 %in% genes.Perseus.AvsV.1000) & (genes.MSqRob.AvsV.1000 %in% genes.Hurdle.AvsV.1000))
# 405

# Control:
length(genes.MSqRob.AvsV.1000)
174+363+58+405

length(genes.Hurdle.AvsV.1000)
134+363+98+405

length(genes.Perseus.AvsV.1000)
439+58+98+405
```

