---
title: "Supplementary tables"
author: "Ludger Goeminne"
date: "09/07/2019"
output: html_document
---

# Supplementary Tables

This document will allow you to reproduce all supplementary tables.

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

Important: make sure you have run the analysis for the missing values in PRIDE (see `1_plots.Rmd`) and the HEART analyse, or load the necessary files as follows:

```{r}
load(paste0(wd,"/save_files_missingness_PRIDE/missingness_PRIDE.RData"))
load(paste0(wd, "/save_files_PXD006675/peptides_HEART1.RData"))
load(paste0(wd, "/save_files_PXD006675/peptides_HEART2.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_Hurdle.RData"))
load(paste0(wd, "/save_files_PXD006675/Perseus_AvsV.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_full.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_co_full.RData"))
```

## 1. Data for the missing values in PRIDE (Supplementary Table 1)

```{r}

# Number of missing values
n_missing

# Number of observed values
n_observed

# Total number of values
n_tot

# % missing values
perc_missingness
```

## 2. Analysis of the HEART dataset with Hurdle, Perseus, MSqRob and the quasi-binomial model (Supplementary Table 2)

```{r}
# Hurdle model
res.HEART.Hurdle[res.HEART.Hurdle$contrast == "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3",]
# openxlsx::write.xlsx(res.HEART.Hurdle[res.HEART.Hurdle$contrast == "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3",], file = "supp_table2_Hurdle.xlsx")

# Perseus with imputation and SAM
Perseus.AvsV
# openxlsx::write.xlsx(Perseus.AvsV, file = "supp_table2_Perseus.xlsx")

# MSqRob
res.HEART.full[res.HEART.full$contrast == "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3",]
# openxlsx::write.xlsx(res.HEART.full[res.HEART.full$contrast == "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3",], file = "supp_table2_MSqRob.xlsx")

# Quasi-binomial model
res.HEART.co.full[res.HEART.co.full$contrast == "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3",]
# openxlsx::write.xlsx(res.HEART.co.full[res.HEART.co.full$contrast == "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3",], file = "supp_table2_quasi-binomial.xlsx")
```

## 3. Compare the left to the right atrium with the hurdle model (Supplementary Table 3)

```{r}
res.HEART.Hurdle[res.HEART.Hurdle$contrast == "regionLA-regionRA",]
# openxlsx::write.xlsx(res.HEART.Hurdle[res.HEART.Hurdle$contrast == "regionLA-regionRA",], file = "supp_table3.xlsx")
```

## 4. Compare the left to the right atrium with the approach of Doll *et al.* (Supplementary Table 4)

```{r}
# SERINC3 and PNMA1 are very likely false positives. PNMA1 is only seen by one peptide in the right atrium, and is not observerd in the left atrium. SERINC3 is also only seen by one peptide in the right atrium, and has one additional identification of that peptide in sample LA4. However, there are many proteins in the dataset like that, which are not declared significant. Overall, the evidence for these two proteins to be differentially abundant, is very poor.

SERINC3 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "SERINC3",]
# Log2 intensities before preprocessing:
log2(SERINC3)
# Log2 intensities after preprocessing:
SERINC3 <- Biobase::exprs(peptides.HEART2)[fData(peptides.HEART2)$gene.name == "SERINC3",]
SERINC3

# boxplot(SERINC3[,c("LA3", "LA4", "LA8", "RA3", "RA4", "RA8", "SepA3", "SepA4", "SepA8", "LV3", "LV4", "LV8", "RV3", "RV4", "RV8", "SepV3", "SepV4", "SepV8")])


PNMA1 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "PNMA1",]
# Log2 intensities before preprocessing:
log2(PNMA1)
# Log2 intensities after preprocessing:
PNMA1 <- Biobase::exprs(peptides.HEART2)[fData(peptides.HEART2)$gene.name == "PNMA1",]
PNMA1

# boxplot(PNMA1[,c("LA3", "LA4", "LA8", "RA3", "RA4", "RA8", "SepA3", "SepA4", "SepA8", "LV3", "LV4", "LV8", "RV3", "RV4", "RV8", "SepV3", "SepV4", "SepV8")])

### Other proteins that are ALSO only seen in one condition, but were not declared significant by Doll et al. and that hold at least as much evidence for DA as the previous two proteins. ###

### ACSM2A, PDE7B, PNPLA7 ###

ACSM2A <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "ACSM2A",]
# Log2 intensities before preprocessing:
log2(ACSM2A)
# Log2 intensities after preprocessing:
ACSM2A <- Biobase::exprs(peptides.HEART2)[fData(peptides.HEART2)$gene.name == "ACSM2A",]
ACSM2A

# boxplot(ACSM2A[,c("LA3", "LA4", "LA8", "RA3", "RA4", "RA8", "SepA3", "SepA4", "SepA8", "LV3", "LV4", "LV8", "RV3", "RV4", "RV8", "SepV3", "SepV4", "SepV8")])

PDE7B <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "PDE7B",]
# Log2 intensities before preprocessing:
log2(PDE7B)
# Log2 intensities after preprocessing:
PDE7B <- Biobase::exprs(peptides.HEART2)[fData(peptides.HEART2)$gene.name == "PDE7B",]
PDE7B

# boxplot(PDE7B[,c("LA3", "LA4", "LA8", "RA3", "RA4", "RA8", "SepA3", "SepA4", "SepA8", "LV3", "LV4", "LV8", "RV3", "RV4", "RV8", "SepV3", "SepV4", "SepV8")])


# A protein similar to SERINC3, where peptide "ILGSLQQGPVTGHQLGLPTEGSK" is seen three times in the right atrium and only once in the LA4 sample:

PNPLA7 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "PNPLA7",]
# Log2 intensities before preprocessing:
log2(PNPLA7)
# Log2 intensities after preprocessing:
PNPLA7 <- Biobase::exprs(peptides.HEART2)[fData(peptides.HEART2)$gene.name == "PNPLA7",]
PNPLA7

# boxplot(PNPLA7[,c("LA3", "LA4", "LA8", "RA3", "RA4", "RA8", "SepA3", "SepA4", "SepA8", "LV3", "LV4", "LV8", "RV3", "RV4", "RV8", "SepV3", "SepV4", "SepV8")])
```
