---
title: "Count analysis for the HEART dataset"
author: "Ludger Goeminne"
date: "22/3/2018"
output: html_document
---

# The quasibinomial model for the HEART dataset

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
```

Important: make sure you have run 1_analysis_HEART_MSqRob.Rmd OR load `peptides.HEART2`, `prot.HEART.co` and `res.HEART.Base` via:

```{r}
load(paste0(wd, "/save_files_PXD006675/peptides_HEART2.RData"))
load(paste0(wd, "/save_files_PXD006675/prot_HEART_co.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_Base.RData"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

## Fit the generalized linear regression with quasibinomial family and logit link function and inspect the results

```{r eval=TRUE}
form.co <- ~ region + subject # + lodds
contrasts <- makeContrast(contrasts=c(
             "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3",
             "regionLA-regionRA",
             "regionLV-regionRV"),
             levels=paste0("region", unique(pData(peptides.HEART2)$region)))


# options(future.globals.maxSize= Inf)

# res.HEART.co <- do_glm(formula = form.co, msnset = prot.HEART.co, group_var = gene.name, contrasts = contrasts, contFun = "contEst") ## 22 min.
# OR: load the model:
load(paste0(wd, "/save_files_PXD006675/res_HEART_co.RData"))

res.HEART.co.full <- res.HEART.Base %>% left_join(res.HEART.co$result)
res.HEART.co.full <- res.HEART.co.full %>% arrange(pvalue)

# Or load the MSqRob result object via:
# load(paste0(wd, "/save_files_PXD006675/res_HEART_co_full.RData"))

# Optional: remove res.HEART.co to free up memory space
rm(res.HEART.co)
gc()
```
