---
title: "Hurdle analysis for the HEART dataset"
author: "Ludger Goeminne"
date: "22/3/2018"
output: html_document
---

# The hurdle model for the HEART dataset

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
```

Important: make sure you have run 1_analysis_PXD006675_MSqRob.Rmd and 2_analysis_PXD006675_counts.Rmd OR load `peptides.HEART2`, `res.HEART.Base`, `res.HEART.full` and `res.HEART.co.full` via:

```{r}
load(paste0(wd, "/save_files_PXD006675/peptides_HEART2.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_Base.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_full.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_co_full.RData"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

The Hurdle model combines the power of MSqRob (without imputation) with the binomial model.

## 1. Join everything together

```{r}
# Join everything together

base.vars <- c("protein", "gene.name", "protein.name", "contrast")

# Set everything in the same order
res.HEART.mm <- full_join(res.HEART.Base,
                  res.HEART.full
                  )

res.HEART.co <- full_join(res.HEART.Base,
                  res.HEART.co.full
                  )
```

## 2. Combine the p-values in the hurdle approach

```{r}
# Combine p-values

z.mm <- -qnorm(res.HEART.mm$pvalue/2)*sign(res.HEART.mm$logFC)
z.co <- -qnorm(res.HEART.co$pvalue/2)*sign(res.HEART.co$logOR)

res.HEART.Hurdle <- res.HEART.Base

res.HEART.Hurdle$logFC <- res.HEART.mm$logFC
res.HEART.Hurdle$logOR <- res.HEART.co$logOR

res.HEART.Hurdle$chisq <- cbind(z.mm^2, z.co^2) %>% rowSums(na.rm = TRUE)
res.HEART.Hurdle$pchisq <- 1 - pchisq(res.HEART.Hurdle$chisq, df = (!is.na(z.mm))+(!is.na(z.co)))

res.HEART.Hurdle <- res.HEART.Hurdle %>% group_by(contrast) %>% mutate(qchisq = pchisq %>%  p.adjust(method = "BH")) %>% ungroup()
res.HEART.Hurdle <- res.HEART.Hurdle %>% arrange(pchisq)

# OR: load the model:
# load(paste0(wd, "/save_files_PXD006675/res_HEART_Hurdle.RData"))

# Optional: remove res.HEART.mm and res.HEART.co to save a little bit of memory space
rm(res.HEART.mm, res.HEART.co)
```

