---
title: "Evaluation of the HEART dataset"
author: "Ludger Goeminne"
date: "03/07/2019"
output: html_document
---

# Evaluation of the different methods on the HEART dataset

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
```

Important: make sure you have run 1_analysis_PXD006675_MSqRob.Rmd and 3_analysis_PXD006675_Hurdle.Rmd OR load `peptides.HEART1`, `peptides.HEART2`, `res.HEART.full` and `res.HEART.Hurdle` via:

```{r}
load(paste0(wd, "/save_files_PXD006675/peptides_HEART1.RData"))
load(paste0(wd, "/save_files_PXD006675/peptides_HEART2.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_full.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_co_full.RData"))
load(paste0(wd, "/save_files_PXD006675/res_HEART_Hurdle.RData"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

## 1. Import the data

### 1.1. Import and preprocess Perseus results

```{r}

### Import Perseus results ###

Perseus.AvsV <- read_xlsx(paste0(wd,"/datasets/PXD006675/41467_2017_1747_MOESM7_ESM.xlsx"), skip = 2)

Perseus.AvsV$`Gene names`[which(grepl("\\\"", Perseus.AvsV$`Gene names`))] <- Perseus.AvsV$`Gene names`[which(grepl("\\\"", Perseus.AvsV$`Gene names`))] %>% gsub(pattern = "\\\"", replacement = "")

######

### Add missing gene annotations ###

Perseus.AvsV[is.na(Perseus.AvsV$`Gene names`),]$`Majority protein IDs`

source(paste0(wd,"/datasets/PXD006675/gene_names.R"))

Perseus.AvsV$`Gene names` <- Perseus.AvsV$`Gene names` %>% as.character
Perseus.AvsV$`Protein names` <- Perseus.AvsV$`Protein names` %>% as.character

for(i in 1:length(proteins)){
  if(proteins[i] %in% Perseus.AvsV$`Majority protein IDs`){
  Perseus.AvsV[Perseus.AvsV$`Majority protein IDs` == proteins[i],]$`Gene names` <- gene.names[i]
  Perseus.AvsV[Perseus.AvsV$`Majority protein IDs` == proteins[i],]$`Protein names` <- protein.names[i]
  }
}

Perseus.AvsV[is.na(Perseus.AvsV$`Gene names`),]
which(Perseus.AvsV$`Gene names` == "")

#######

### Arrange Perseus results according to the results of the SAM test with s0 = 0.1, as applied by Doll et al.###

pval <- 10^(-Perseus.AvsV$`minusLOG(P-value)`)

Tval <- -qt(pval/2, 16)*sign(Perseus.AvsV$Difference)

SE <- Perseus.AvsV$Difference/Tval

TvalS0 <- Perseus.AvsV$Difference/(SE+0.1)

Perseus.AvsV$TvalS0 <- TvalS0

Perseus.AvsV <- Perseus.AvsV %>% arrange(desc(abs(TvalS0)))

# Check: the first 1629 are the significant ones, thus we ranked correctly!
sum(!is.na(Perseus.AvsV$`Significantly altered proteins (ventricular vs. atrial region)`))
# 1629
all(which(!is.na(Perseus.AvsV$`Significantly altered proteins (ventricular vs. atrial region)`)) == 1:1629)
# TRUE

# Or load Perseus.AvsV via:
# load(paste0(wd, "/save_files_PXD006675/Perseus_AvsV.RData"))
#######
```

## 1.2. Import the proteinGroups.txt file in R

With their 379 and 405 MB respectively, the peptides.txt file and proteinGroups.txt files are too big to put on our GitHub server. However, they are freely available from https://www.ebi.ac.uk/pride/archive/projects/PXD006675/files. Unzip the search.zip file and look for these files in the "txt" folder.

**Option 1 for importing the data: **
download the peptides.txt and proteinGroups.txt files from the PRIDE repository and put them in the folder "datasets" under a subfolder called "PXD006675". Then, execute the following block of code:

```{r eval = FALSE}
file.proteinGroups <- paste0(wd,"/datasets/PXD006675/proteinGroups.txt")

exprs_col <- grepEcols(file.proteinGroups, "LFQ intensity ", split = "\t")
proteinGroups.HEART <- readMSnSet2(file.proteinGroups, ecol = exprs_col, fnames = c("Protein.IDs"), sep = "\t")
sampleNames(proteinGroups.HEART) <- str_replace(sampleNames(proteinGroups.HEART), "LFQ.intensity.", "") %>% make.names

# We only keep the samples in which we are interested, we also log2-transform the data
proteinGroups.HEART <- MSnSet(exprs = Biobase::exprs(proteinGroups.HEART)[,c("RA3", "RA4", "RA8", "LA3", "LA4", "LA8", "SepA3", "SepA4", "SepA8", "RV3", "RV4", "RV8", "LV3", "LV4", "LV8", "SepV3", "SepV4", "SepV8")] %>% log2, fData = AnnotatedDataFrame(fData(proteinGroups.HEART)[,c("Gene.names", "Protein.IDs", "Majority.protein.IDs", "Protein.names")]), pData = AnnotatedDataFrame(pData(proteinGroups.HEART)[c("RA3", "RA4", "RA8", "LA3", "LA4", "LA8", "SepA3", "SepA4", "SepA8", "RV3", "RV4", "RV8", "LV3", "LV4", "LV8", "SepV3", "SepV4", "SepV8"),]))

Biobase::exprs(proteinGroups.HEART)[is.infinite(Biobase::exprs(proteinGroups.HEART))] <- NA

# Check: all Perseus protein IDs are present in the proteinGroups file (as it should be).
all(Perseus.AvsV$`Protein IDs` %in% fData(proteinGroups.HEART)$`Protein.IDs`)
# TRUE

# Now remove the proteins that were filtered out by Doll et al.
proteinGroups.HEART1 <- proteinGroups.HEART[fData(proteinGroups.HEART)$`Protein.IDs` %in% Perseus.AvsV$`Protein IDs` %>% which]

# Remove proteinGroups.HEART to save memory space
rm(proteinGroups.HEART)
gc()
```

**Option 2:** load the `proteinGroups.HEART1` MSnSet object:

```{r}
load(paste0(wd,"/save_files_PXD006675/proteinGroups_HEART1.RData"))
```

## 1.3. Export the filtered data to a tab-delimited file

```{r}
df.proteinGroups.HEART <- cbind(Biobase::fData(proteinGroups.HEART1), Biobase::exprs(proteinGroups.HEART1))
colnames(df.proteinGroups.HEART)[colnames(df.proteinGroups.HEART) %in% c("Gene.names", "Protein.IDs", "Majority.protein.IDs", "Protein.names")] <- c("Gene names", "Protein IDs", "Majority protein IDs", "Protein names")

# df.proteinGroups.HEART %>% write.table(file = paste0(wd, "/datasets/PXD006675/proteinGroups_HEART_filtered.txt"), quote = FALSE, sep = "\t", row.names = FALSE)
# This file is already provided in the /datasets/PXD006675 folder.
```

## 2. The effect of imputation with Perseus

Here, we investigate the effect of Perseus imputation. We perform 5 analyses where we impute the data. We also do the analysis without imputation.

### 2.1. Repeated imputation with Perseus

For the Perseus analysis, we carried out the following steps in Perseus 1.6.0.7, as described in Doll *et al.*
(https://www.nature.com/articles/s41467-017-01747-2).

1. Generic matrix upload of `".../datasets/PXD006675/proteinGroups_HEART_filtered.txt"`. All data columns (`RA3`, `RA4`, ..., `SepV8`) are imported as `Main` columns. `Gene names`, `Protein IDs`, `Majority Protein IDs` and `Protein names` are imported as `Text` columns.

2. Annotate the columns according to `heart region` (i.e. LA, LV, RA, RV, SepA and SepV) via `"Annot. rows -> Categorical annotation rows"`.

3. Annotate the columns according to `heart chamber` (i.e. `atrium` and `ventricle`) via `"Annot. rows -> Categorical annotation rows"`.

4. We do (a.) five times an imputation via `"Imputation -> Replace missing values from normal distribtion"` with `(width = 0.15, down shift = 1.8)` and Mode `Separately for each column` or (b.) no imputation at all.

5. For both the imputed (7.a.) data and not imputed data (7.b.), we do unpaired two-sample t-tests via `"Tests -> Two samples test"` for (atrium vs. ventricle). We thus specify the following in Perseus:

`First group (right):`
`atrium`

`Second group (left):`
`ventricle`

We do a default Permutation-based FDR with a default cut-off at 0.05. We set S0 to 0.1.
We do the default 250 randomizations. We do not -log10 transform the p-values.

We saved these files in the `simulations_PXD006675_Perseus` folder.
`imp_result1.txt` until `imp_result5.txt` are the analyses where the missing values were imputed.
`no_imp_result.txt` is the analysis where the missing values were not imputed.

We import these results and sort them by significance.

```{r}
### Function to arrange the results according to the results of the SAM test with s0 = 0.1, as applied by Doll et al. ###
arrange_SAM <- function(df, int_cols = 1:18){
  pval <- df$`Student's T-test p-value`
  dfs <- (df[, int_cols] %>%  is.na %>% `!` %>% rowSums)-2
  Tval <- -qt(pval/2, dfs)*sign(df$`Student's T-test Difference atrium_ventricle`)
  SE <- df$`Student's T-test Difference atrium_ventricle`/Tval
  df$TvalS0 <- df$`Student's T-test Difference atrium_ventricle`/(SE+0.1)
  df <- df %>% arrange(desc(abs(df$TvalS0)))
  df$pvalS0 <- (df$TvalS0 %>% abs %>% pt(df=dfs, lower=FALSE))*2
  df <- df %>% arrange(df$pvalS0)
  df <- df %>% arrange(`Student's T-test q-value atrium_ventricle`)
  df$pvalS0 <- NULL # The degrees of freedom are not completely correct for SAM, we only used this metric for ranking. This is needed only for the SAM simulation with non-imputed data, where the degrees of freedom can differ from protein to protein due to missing values. Therefore, it is best to remove it again to not give a wrong impression about the real underlying p-value.
  return(df)
}
```

Import the result from the 5 imputation analyses and the one analysis without imputation into R.

```{r}
colnames <- read.table(paste0(wd, "/simulations_PXD006675_Perseus/imp_result1.txt"), header = FALSE,
                   sep = "\t", quote = "", comment.char = "", nrow = 1)

### Imputed results 1 ###

imp_result1 <- read.table(paste0(wd, "/simulations_PXD006675_Perseus/imp_result1.txt"), header = FALSE,
                   sep = "\t", quote = "", comment.char = "", skip = 4)
colnames(imp_result1) <- unlist(colnames)
imp_result1 <- arrange_SAM(imp_result1)

# Check: the first 1522 are the significant ones, thus we ranked correctly!
sum(imp_result1$`Student's T-test Significant atrium_ventricle`=="+")
# 1522
all(which(imp_result1$`Student's T-test Significant atrium_ventricle` == "+") == 1:1522)
# TRUE

### Imputed results 2 ###

imp_result2 <- read.table(paste0(wd, "/simulations_PXD006675_Perseus/imp_result2.txt"), header = FALSE,
                    sep = "\t", quote = "", comment.char = "", skip = 4)
colnames(imp_result2) <- unlist(colnames)
imp_result2 <- arrange_SAM(imp_result2)

# Check: the first 1524 are the significant ones, thus we ranked correctly!
sum(imp_result2$`Student's T-test Significant atrium_ventricle`=="+")
# 1524
all(which(imp_result2$`Student's T-test Significant atrium_ventricle` == "+") == 1:1524)
# TRUE

### Imputed results 3 ###

imp_result3 <- read.table(paste0(wd, "/simulations_PXD006675_Perseus/imp_result3.txt"), header = FALSE,
                    sep = "\t", quote = "", comment.char = "", skip = 4)
colnames(imp_result3) <- unlist(colnames)
imp_result3 <- arrange_SAM(imp_result3)

# Check: the first 1535 are the significant ones, thus we ranked correctly!
sum(imp_result3$`Student's T-test Significant atrium_ventricle`=="+")
# 1535
all(which(imp_result3$`Student's T-test Significant atrium_ventricle` == "+") == 1:1535)
# TRUE

### Imputed results 4 ###

imp_result4 <- read.table(paste0(wd, "/simulations_PXD006675_Perseus/imp_result4.txt"), header = FALSE,
                    sep = "\t", quote = "", comment.char = "", skip = 4)
colnames(imp_result4) <- unlist(colnames)
imp_result4 <- arrange_SAM(imp_result4)

# Check: the first 1530 are the significant ones, thus we ranked correctly!
sum(imp_result4$`Student's T-test Significant atrium_ventricle`=="+")
# 1530
all(which(imp_result4$`Student's T-test Significant atrium_ventricle` == "+") == 1:1530)
# TRUE

### Imputed results 5 ###

imp_result5 <- read.table(paste0(wd, "/simulations_PXD006675_Perseus/imp_result5.txt"), header = FALSE,
                    sep = "\t", quote = "", comment.char = "", skip = 4)
colnames(imp_result5) <- unlist(colnames)
imp_result5 <- arrange_SAM(imp_result5)

# Check: the first 1541 are the significant ones, thus we ranked correctly!
sum(imp_result5$`Student's T-test Significant atrium_ventricle`=="+")
# 1541
all(which(imp_result5$`Student's T-test Significant atrium_ventricle` == "+") == 1:1541)
# TRUE

### Not imputed results ###

noimp_result <- read.table(paste0(wd, "/simulations_PXD006675_Perseus/noimp_result.txt"), header = FALSE,
                    sep = "\t", quote = "", comment.char = "", skip = 4)
colnames(noimp_result) <- unlist(colnames)
noimp_result <- arrange_SAM(noimp_result)

# Check: the first 1231 are the significant ones, thus we ranked correctly!
sum(noimp_result$`Student's T-test Significant atrium_ventricle`=="+")
# 1231
all(which(noimp_result$`Student's T-test Significant atrium_ventricle` == "+") == 1:1231)
# TRUE
```

Even though our repeated imputation simulations come close to the results of Doll *et al.*, we still declare around 100 proteins less as differentially abundant on the 5% FDR level. It is difficult to speculate about the cause for this, as we used the same subset as the authors used in their original publication. However, by comparing our simulations, we can get an idea of the impact of imputing data from analysis to analysis.

```{r}
sum(imp_result1$`Student's T-test Significant atrium_ventricle`=="+")
# 1522
sum(imp_result2$`Student's T-test Significant atrium_ventricle`=="+")
# 1524
sum(imp_result3$`Student's T-test Significant atrium_ventricle`=="+")
# 1535
sum(imp_result4$`Student's T-test Significant atrium_ventricle`=="+")
# 1530
sum(imp_result5$`Student's T-test Significant atrium_ventricle`=="+")
# 1541

sum(!is.na(Perseus.AvsV$`Significantly altered proteins (ventricular vs. atrial region)`))
# 1629
```

### 2.2. Overlap in the first 1500 proteins

```{r}
N0 <- noimp_result$`Protein IDs`[1:1500]
N1 <- imp_result1$`Protein IDs`[1:1500]
N2 <- imp_result2$`Protein IDs`[1:1500]
N3 <- imp_result3$`Protein IDs`[1:1500]
N4 <- imp_result4$`Protein IDs`[1:1500]
N5 <- imp_result5$`Protein IDs`[1:1500]

1-sum(N1 %in% N2)/1500
1-sum(N1 %in% N3)/1500
1-sum(N1 %in% N4)/1500
1-sum(N1 %in% N5)/1500

1-sum(N2 %in% N1)/1500
1-sum(N2 %in% N3)/1500
1-sum(N2 %in% N4)/1500
1-sum(N2 %in% N5)/1500

1-sum(N3 %in% N2)/1500
1-sum(N3 %in% N1)/1500
1-sum(N3 %in% N4)/1500
1-sum(N3 %in% N5)/1500

1-sum(N4 %in% N2)/1500
1-sum(N4 %in% N3)/1500
1-sum(N4 %in% N1)/1500
1-sum(N4 %in% N5)/1500

1-sum(N5 %in% N2)/1500
1-sum(N5 %in% N3)/1500
1-sum(N5 %in% N4)/1500
1-sum(N5 %in% N1)/1500
# -> around 4% to 5%

1-sum(N0 %in% N1)/1500
1-sum(N0 %in% N2)/1500
1-sum(N0 %in% N3)/1500
1-sum(N0 %in% N4)/1500
1-sum(N0 %in% N5)/1500
# -> around 22%
```

## 3. Hurdle's results on the comparison of the atrial versus ventricular regions

### 3.1. Select hurdle, MSqRob and quasibinomial results for this comparison

```{r}
# Hurdle results for the atrial vs. ventricular comparison
hurdle.AvsV <- res.HEART.Hurdle[res.HEART.Hurdle$contrast == "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3",]

# MSqRob results for the atrial vs. ventricular comparison
MSqRob.AvsV <- res.HEART.full[res.HEART.full$contrast == "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3",]

# Quasibinomial results for the atrial vs. ventricular comparison
qBinom.AvsV <- res.HEART.co.full[res.HEART.co.full$contrast == "(regionLA+regionRA+regionSepA)/3-(regionLV+regionRV+regionSepV)/3",]
```

### 3.2. Apply the two-stage procedure

For the two-stage procedure, we make use of the stageR package.
The latest version of stageR can be found at: https://github.com/statOmics/stageR.
We made use of stageR version 1.3.29.

```{r}
# First stage: hurdle's p-values
pScreen <- hurdle.AvsV %>% arrange(gene.name) %>% pull(pchisq)
names(pScreen) <- hurdle.AvsV %>% arrange(gene.name) %>% pull(gene.name)

# Second stage: the p-values of the MSqRob and the quasibinomial component
pConfirmation <- data.frame(
p_MSqRob <- MSqRob.AvsV %>% arrange(gene.name) %>% pull(pvalue),
p_qBinom <- qBinom.AvsV %>% arrange(gene.name) %>% pull(pvalue)
) %>% as.matrix

rownames(pConfirmation) <- MSqRob.AvsV %>% arrange(gene.name) %>% pull(gene.name)

# Check
all(MSqRob.AvsV %>% arrange(gene.name) %>% pull(gene.name) == qBinom.AvsV %>% arrange(gene.name) %>% pull(gene.name))
# TRUE
all(hurdle.AvsV %>% arrange(gene.name) %>% pull(gene.name) == qBinom.AvsV %>% arrange(gene.name) %>% pull(gene.name))
# TRUE

# Create a stageR object
stageRObj <- stageR(pScreen = pScreen, pConfirmation = pConfirmation, pScreenAdjusted = FALSE)
# We don't need to correct the FWER in our second stage because signficance in the first stage already guarantees that max. one null hypothesis is incorrectly rejected. We set the significance threshold at 5% FDR.
stageRObj <- stageWiseAdjustment(object = stageRObj, method = "none", alpha = 0.05, allowNA = TRUE)

# Inspect the stageR object
head(getAdjustedPValues(stageRObj, onlySignificantGenes = FALSE, order = FALSE))

res <- getResults(stageRObj)
head(res)
tail(res)
res[is.na(res)] <- 0

# Check
sum(res[,1])
# 1765
sum(hurdle.AvsV$qchisq < 0.05, na.rm = TRUE)
# 1765 gene identifiers were significant at the 5% FDR level in the first stage

table(rowSums(res[,c(2,3)]))
# 0    1    2
# 113  1212  440

1212+440
# 1652 gene identifiers are significant in at least one of both confirmation tests.
1765-1652
# For 113 gene identifiers, significance in the first stage could not clearly be attributed to either DD or DA

# Check:
113+1212+440
# 1765

# 440 gene identifiers were both DA and DD in the second stage

colSums(res, na.rm = TRUE)
# padjScreen   p_MSqRob   p_qBinom
# 1765       1255      837

1255-440
# 815 proteins were DA, but not DD in the second stage

837-440
# 397 gene identifiers were DD, but not DA in the second stage
```

### 3.3. Direction of the OR and FC estimates that are both DD and DA in stage 2

What is the direction of the OR and FC estimates for the proteins that are both DD and DA?

```{r}
bothDD_DA <- hurdle.AvsV[hurdle.AvsV$gene.name %in% (res[rowSums(res[,c(2,3)])==2,] %>% rownames),]
dim(bothDD_DA)
# 440   8

(bothDD_DA$logFC %>% sign == bothDD_DA$logOR %>% sign) %>% sum
# 440

440/440
# 100% => all FC and OR estimates in the same direction for these gene identifiers
```

## 4. Comparison with the original paper

### 4.1. Select overlapping gene identifiers for each approach

```{r}
### We don't want to assess the influence of different preprocessing pipelines, thus, we only assess the overlapping genes. ###

# B., C. genes in the Perseus preprocessing, but not in Hurdle/MSqRob preprocessing
not.in.Hurdle <- Perseus.AvsV$`Gene names`[!(Perseus.AvsV$`Gene names` %in% hurdle.AvsV$gene.name)]

# B., C. genes in the Hurdle/MSqRob preprocessing, but not in Perseus preprocessing
not.in.Perseus <- hurdle.AvsV$gene.name[!(hurdle.AvsV$gene.name %in% Perseus.AvsV$`Gene names`)]

Perseus.AvsV.ov <- Perseus.AvsV[!(Perseus.AvsV$`Gene names` %in% not.in.Hurdle),]
hurdle.AvsV.ov <- hurdle.AvsV[!(hurdle.AvsV$gene.name %in% not.in.Perseus),]
MSqRob.AvsV.ov <- MSqRob.AvsV[!(MSqRob.AvsV$gene.name %in% not.in.Perseus),]

# Note that Perseus has multiple entries for the same gene name, as we grouped per gene name and they grouped per protein ID.
length(unique(Perseus.AvsV.ov$`Gene names`))
length(unique(hurdle.AvsV.ov$gene.name))
length(unique(MSqRob.AvsV.ov$gene.name))
# 7822

# Or: load these files via:
# load(paste0(wd, "/save_files_PXD006675/overlap_PHM.RData"))
```

### 4.2. Left vs. right ventricle & left vs. right atrium

```{r}

res.HEART.Hurdle[res.HEART.Hurdle$contrast == "regionLV-regionRV",]
# No significant hits, just like the authors

# Select gene identifiers for our analysis of atria vs. ventricles
res.HEART.Hurdle[res.HEART.Hurdle$contrast == "regionLA-regionRA",]

# SERINC3 and PNMA1 are very likely false positives. PNMA1 is only seen by one peptide in the right atrium, and is not observerd in the left atrium. SERINC3 is also only seen by one peptide in the right atrium, and has one additional identification of that peptide in sample LA4. However, there are many proteins in the dataset like that, which are not declared significant. Overall, the evidence for these two proteins to be differentially abundant, is very poor.

SERINC3 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "SERINC3",]
sum(SERINC3[,c("LA3", "LA4", "LA8")]) - sum(SERINC3[,c("RA3", "RA4", "RA8")])
# -3574222000

SERINC3_1 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "SERINC3",] %>% log2
SERINC3_1 <- as_tibble(SERINC3_1) %>% mutate(Sequence = row.names(SERINC3_1))

SERINC3_1 <- SERINC3_1 %>% gather(key = "sample", value = log2_intensity_before_preprocessing, -Sequence)
SERINC3_1 <- SERINC3_1[is.finite(SERINC3_1$log2_intensity_before_preprocessing),]

SERINC3_2 <- Biobase::exprs(peptides.HEART2)[fData(peptides.HEART2)$gene.name == "SERINC3",]
SERINC3_2 <- as_tibble(SERINC3_2) %>% mutate(Sequence = row.names(SERINC3_2))

SERINC3_2 <- SERINC3_2 %>% gather(key = "sample", value = log2_intensity_after_preprocessing, -Sequence)
SERINC3_2 <- SERINC3_2[is.finite(SERINC3_2$log2_intensity_after_preprocessing),]

merge(SERINC3_1, SERINC3_2)

PNMA1 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "PNMA1",]
sum(PNMA1[,c("LA3", "LA4", "LA8")]) - sum(PNMA1[,c("RA3", "RA4", "RA8")])
# -1321840000

PNMA1_1 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "PNMA1",] %>% log2
PNMA1_1 <- as_tibble(PNMA1_1) %>% mutate(Sequence = row.names(PNMA1_1))

PNMA1_1 <- PNMA1_1 %>% gather(key = "sample", value = log2_intensity_before_preprocessing, -Sequence)
PNMA1_1 <- PNMA1_1[is.finite(PNMA1_1$log2_intensity_before_preprocessing),]

PNMA1_2 <- Biobase::exprs(peptides.HEART2)[fData(peptides.HEART2)$gene.name == "PNMA1",]
PNMA1_2 <- as_tibble(PNMA1_2) %>% mutate(Sequence = row.names(PNMA1_2))

PNMA1_2 <- PNMA1_2 %>% gather(key = "sample", value = log2_intensity_after_preprocessing, -Sequence)
PNMA1_2 <- PNMA1_2[is.finite(PNMA1_2$log2_intensity_after_preprocessing),]

merge(PNMA1_1, PNMA1_2)

# Other proteins that are ALSO only seen in one condition

# ACSM2A, IGLV5-45, PDE7B

ACSM2A <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "ACSM2A",]
sum(ACSM2A[,c("LA3", "LA4", "LA8")]) - sum(ACSM2A[,c("RA3", "RA4", "RA8")])
# 110030000

ACSM2A_1 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "ACSM2A",] %>% log2
ACSM2A_1 <- as_tibble(ACSM2A_1) %>% mutate(Sequence = row.names(ACSM2A_1))

ACSM2A_1 <- ACSM2A_1 %>% gather(key = "sample", value = log2_intensity_before_preprocessing, -Sequence)
ACSM2A_1 <- ACSM2A_1[is.finite(ACSM2A_1$log2_intensity_before_preprocessing),]

ACSM2A_2 <- Biobase::exprs(peptides.HEART2)[fData(peptides.HEART2)$gene.name == "ACSM2A",]
ACSM2A_2 <- as_tibble(ACSM2A_2) %>% mutate(Sequence = row.names(ACSM2A_2))

ACSM2A_2 <- ACSM2A_2 %>% gather(key = "sample", value = log2_intensity_after_preprocessing, -Sequence)
ACSM2A_2 <- ACSM2A_2[is.finite(ACSM2A_2$log2_intensity_after_preprocessing),]

merge(ACSM2A_1, ACSM2A_2)

PDE7B <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "PDE7B",]
sum(PDE7B[,c("LA3", "LA4", "LA8")]) - sum(PDE7B[,c("RA3", "RA4", "RA8")])
# 186101600

PDE7B_1 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "PDE7B",] %>% log2
PDE7B_1 <- as_tibble(PDE7B_1) %>% mutate(Sequence = row.names(PDE7B_1))

PDE7B_1 <- PDE7B_1 %>% gather(key = "sample", value = log2_intensity_before_preprocessing, -Sequence)
PDE7B_1 <- PDE7B_1[is.finite(PDE7B_1$log2_intensity_before_preprocessing),]

PDE7B_2 <- Biobase::exprs(peptides.HEART2)[fData(peptides.HEART2)$gene.name == "PDE7B",]
PDE7B_2 <- as_tibble(PDE7B_2) %>% mutate(Sequence = row.names(PDE7B_2))

PDE7B_2 <- PDE7B_2 %>% gather(key = "sample", value = log2_intensity_after_preprocessing, -Sequence)
PDE7B_2 <- PDE7B_2[is.finite(PDE7B_2$log2_intensity_after_preprocessing),]

merge(PDE7B_1, PDE7B_2)

# A protein similar to SERINC3, where peptide "ILGSLQQGPVTGHQLGLPTEGSK" is seen three times in the right atrium and only once in the LA4 sample:

PNPLA7 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "PNPLA7",]
sum(PNPLA7[,c("LA3", "LA4", "LA8")]) - sum(PNPLA7[,c("RA3", "RA4", "RA8")])
# -322611000

PNPLA7_1 <- Biobase::exprs(peptides.HEART1)[fData(peptides.HEART1)$gene.name == "PNPLA7",] %>% log2
PNPLA7_1 <- as_tibble(PNPLA7_1) %>% mutate(Sequence = row.names(PNPLA7_1))

PNPLA7_1 <- PNPLA7_1 %>% gather(key = "sample", value = log2_intensity_before_preprocessing, -Sequence)
PNPLA7_1 <- PNPLA7_1[is.finite(PNPLA7_1$log2_intensity_before_preprocessing),]

PNPLA7_2 <- Biobase::exprs(peptides.HEART2)[fData(peptides.HEART2)$gene.name == "PNPLA7",]
PNPLA7_2 <- as_tibble(PNPLA7_2) %>% mutate(Sequence = row.names(PNPLA7_2))

PNPLA7_2 <- PNPLA7_2 %>% gather(key = "sample", value = log2_intensity_after_preprocessing, -Sequence)
PNPLA7_2 <- PNPLA7_2[is.finite(PNPLA7_2$log2_intensity_after_preprocessing),]

merge(PNPLA7_1, PNPLA7_2)

# -> none of them are contaminants, reverse or only identified by site
```
