---
title: "MSqRob analysis with mixed imputation for the CPTAC dataset"
author: "Ludger Goeminne"
date: "02/12/2019"
output: html_document
---

# MSqRob with mixed imputation for the CPTAC dataset

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
```

Important: make sure you have run 1_analysis_CPTAC_MSqRob.Rmd OR load `peptides.CPTAC2` and `res.CPTAC.Base` via:

```{r}
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC2.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_Base.RData"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

## 1. Data preprocessing

### 1.1. Mixed imputation as proposed in the DEP vignette

See:
https://bioconductor.org/packages/release/bioc/vignettes/DEP/inst/doc/MissingValues.html#mixed-imputation-on-proteins-rows

```{r eval=TRUE}
### Extra compared to the normal MSqRob approach: impute with mixed imputation

count.peptides <- MSnSet2df(peptides.CPTAC2) %>% as_tibble %>% group_by(feature, condition) %>% summarize(n())
peptides_MNAR <- names(which(table(count.peptides$feature) != 3))

## Get a logical vector
  MNAR <- rownames(exprs(peptides.CPTAC2)) %in% peptides_MNAR
  print('Number of proteins that are Missing Not At Random')
  print(table(MNAR))

 ## Perform a mixed imputation
  # The DEP example in the vignette considers a protein to have missing values not at random (MNAR) if it has missing values in all replicates of at least one condition.
randna <- !MNAR
peptides.CPTAC.mixed <- peptides.CPTAC2
# Note: we need to set colmax to 0.9 because there are too many missing values...
exprs(peptides.CPTAC.mixed)[randna, ] <- exprs(impute(peptides.CPTAC2, "knn", colmax = 0.9))[randna,]
exprs(peptides.CPTAC.mixed)[!randna, ] <- exprs(impute(peptides.CPTAC2, "MinProb"))[!randna,]
```

### 1.2. Evaluate the imputation

We make a histogram to evaluate the imputation. Blue are the original values, red are the imputed values.

```{r eval=TRUE}
p1 <- hist(exprs(peptides.CPTAC.mixed), breaks=25, plot = FALSE)
p2 <- hist(exprs(peptides.CPTAC2), breaks=25, plot = FALSE)
plot(p1, col=rgb(255,0,0, maxColorValue = 255), xlim=c(13,28), border=rgb(255,165,0, maxColorValue = 255), main = "mixed imputation", xlab = "log2(peptide intensity)")  # first histogram
plot(p2, col=rgb(0,0,255, maxColorValue = 255), xlim=c(13,28), border=rgb(0,165,255, maxColorValue = 255), add=TRUE)  # second histogram
```

### 1.3. Inspect the data

We first want to know how many human UPS1 proteins are present in the dataset.

```{r eval=TRUE}
p <- MSnSet2df(peptides.CPTAC.mixed)
select(p, protein) %>% distinct %>% count(ups = grepl('ups', protein))
# # A tibble: 2 x 2
#   ups       n
#   <lgl> <int>
# 1 FALSE  1343
# 2 TRUE     38
```

There are 38 UPS1 proteins present in the dataset.

Here we check how many proteins have only peptides in 1 spike-in condition.

```{r eval=TRUE}
condition1 <- select(p, protein, condition) %>% distinct %>% group_by(protein) %>% count %>% ungroup %>% filter(n == 1)
condition1 %>% count(ups = grepl('ups', protein))
# A tibble: 0 x 2
# ... with 2 variables: ups <lgl>, nn <int>
```

There are no proteins that are only detected in one spike-in condition.

```{r eval=TRUE}
lab1 <- select(p, protein, lab) %>% distinct %>% group_by(protein) %>% count %>% ungroup %>% filter(n == 1)
lab1 %>% count(ups = grepl('ups', protein))
# A tibble: 0 x 2
# ... with 2 variables: protein <fct>, n <int>
rm(condition1, lab1, p, p1, p2)
```

There are no proteins that are only detected in one lab.

## 2. MSqRob with kNN analysis

### 2.1. Drop levels

Since we imputed the data, we don't need any additional filtering steps.

```{r eval=TRUE}
# Drop levels
peptides.CPTAC.mixed <- MSnbase::MSnSet(exprs = Biobase::exprs(peptides.CPTAC.mixed), fData = droplevels(Biobase::fData(peptides.CPTAC.mixed)), pData = Biobase::pData(peptides.CPTAC.mixed))

dim(peptides.CPTAC.mixed)
# 9377   27
length(unique(fData(peptides.CPTAC.mixed)$protein))
# 1381

# Alternative: load the imputed MSnSet object "peptides.CPTAC.mixed" from:
# load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC_mixed.RData"))
```

### 2.2. Specify formula, create contrast matrix

```{r eval=TRUE}
form <- expression ~ (1|condition) + (1|lab) + (1|sample) + (1|sequence)
contrasts <- contrast_helper(form, peptides.CPTAC.mixed, condition)
```

### 2.3. Fit robust mixed models and test contrasts

```{r eval=TRUE}
# Execute the model by calling this command:
# res.CPTAC.mixed <- do_mm(formula = form, msnset = peptides.CPTAC.mixed, group_var = protein, contrasts = contrasts, type_df = "conservative", max_iter = 20) # 8 minutes on our system
# OR: load the model:
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_mixed.RData"))
```

### 2.4. Check results

```{r eval=TRUE}
sum(table(res.CPTAC.mixed$result$protein) == 3)
```

For all 1381 proteins, all contrasts could be estimated (since we imputed the missing values).

### 2.5. Add proteins for which no models could be fitted

```{r eval=TRUE}
res.CPTAC.mixed.full <- res.CPTAC.Base %>% left_join(res.CPTAC.mixed$result)
res.CPTAC.mixed.full <- res.CPTAC.mixed.full %>% arrange(pvalue)

# Or load the MSqRob result object via:
# load(paste0(wd, "/save_files_CPTAC/res_CPTAC_mixed_full.RData"))

# Optional: remove res.CPTAC.mixed to free up memory space
rm(res.CPTAC.mixed)
gc()
```
