---
title: "EdgeR analysis for the CPTAC dataset"
author: "Ludger Goeminne"
date: "7-11-2019"
output: html_document
---

# EdgeR for the CPTAC dataset

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
```

Important: make sure you have run 1_analysis_CPTAC_MSqRob.Rmd and prepare_spectral_counts.Rmd OR load `peptides.CPTAC2`, `prot.CPTAC.co`, `prot.CPTAC.spectral.co` and `res.CPTAC.Base` via:

```{r}
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC2.RData"))
load(paste0(wd, "/save_files_CPTAC/prot_CPTAC_co.RData"))
load(paste0(wd, "/save_files_CPTAC/prot_CPTAC_spectral_co.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_Base.RData"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

## 1. edgeR analysis for peptide counts

```{r}
condition <- pData(prot.CPTAC.co)$condition
lab <- pData(prot.CPTAC.co)$lab
y <- DGEList(counts = exprs(prot.CPTAC.co), group = condition)
y <- calcNormFactors(y)
design <- model.matrix(~condition + lab)
y <- estimateDisp(y, design)
fit <- glmQLFit(y, design)
# B vs. A
qlf.BvsA <- glmQLFTest(fit, contrast=c(0, 1, 0, 0, 0))
res.edgeR.BvsA <- topTags(qlf.BvsA, n = Inf)$table
res.edgeR.BvsA$contrast <- "condition6B-condition6A"
res.edgeR.BvsA <- res.edgeR.BvsA %>% cbind(protein = rownames(.),.)

# C vs. A
qlf.CvsA <- glmQLFTest(fit, contrast=c(0, 0, 1, 0, 0))
res.edgeR.CvsA <- topTags(qlf.CvsA, n = Inf)$table
res.edgeR.CvsA$contrast <- "condition6C-condition6A"
res.edgeR.CvsA <- res.edgeR.CvsA %>% cbind(protein = rownames(.),.)

# C vs. B
qlf.CvsB <- glmQLFTest(fit, contrast=c(0, -1, 1, 0, 0))
res.edgeR.CvsB <- topTags(qlf.CvsB, n = Inf)$table
res.edgeR.CvsB$contrast <- "condition6C-condition6B"
res.edgeR.CvsB <- res.edgeR.CvsB %>% cbind(protein = rownames(.),.)

res.edgeR <- rbind(res.edgeR.BvsA, res.edgeR.CvsA, res.edgeR.CvsB)
res.CPTAC.edgeR.full <- res.CPTAC.Base %>% left_join(res.edgeR)
res.CPTAC.edgeR.full <- res.CPTAC.edgeR.full %>% arrange(PValue)

# Or load the result object via:
# load(paste0(wd, "/save_files_CPTAC/res_CPTAC_edgeR_full.RData"))
```

## 2. edgeR analysis for spectral counts

```{r}
# edgeR analysis on spectral counts
condition <- pData(prot.CPTAC.spectral.co)$condition
lab <- pData(prot.CPTAC.spectral.co)$lab
y <- DGEList(counts = exprs(prot.CPTAC.spectral.co), group = condition)
y <- calcNormFactors(y)
design <- model.matrix(~condition + lab)
y <- estimateDisp(y, design)
fit <- glmQLFit(y, design)
# B vs. A
qlf.BvsA <- glmQLFTest(fit, contrast=c(0, 1, 0, 0, 0))
res.edgeR.BvsA <- topTags(qlf.BvsA, n = Inf)$table
res.edgeR.BvsA$contrast <- "condition6B-condition6A"
res.edgeR.BvsA <- res.edgeR.BvsA %>% cbind(protein = rownames(.),.)

# C vs. A
qlf.CvsA <- glmQLFTest(fit, contrast=c(0, 0, 1, 0, 0))
res.edgeR.CvsA <- topTags(qlf.CvsA, n = Inf)$table
res.edgeR.CvsA$contrast <- "condition6C-condition6A"
res.edgeR.CvsA <- res.edgeR.CvsA %>% cbind(protein = rownames(.),.)

# C vs. B
qlf.CvsB <- glmQLFTest(fit, contrast=c(0, -1, 1, 0, 0))
res.edgeR.CvsB <- topTags(qlf.CvsB, n = Inf)$table
res.edgeR.CvsB$contrast <- "condition6C-condition6B"
res.edgeR.CvsB <- res.edgeR.CvsB %>% cbind(protein = rownames(.),.)

res.edgeR <- rbind(res.edgeR.BvsA, res.edgeR.CvsA, res.edgeR.CvsB)
res.CPTAC.edgeR.sc.full <- res.CPTAC.Base %>% left_join(res.edgeR)
res.CPTAC.edgeR.sc.full <- res.CPTAC.edgeR.sc.full %>% arrange(PValue)

# Or load the result object via:
# load(paste0(wd, "/save_files_CPTAC/res_CPTAC_edgeR_sc_full.RData"))
```



