---
title: "ProDA analysis for the CPTAC dataset"
author: "Ludger Goeminne"
date: "3-12-2019"
output: html_document
---

# Analysis of the Protein Differential Abundance (ProDA) workflow for the CPTAC dataset based on LFQ-summarized data

We base ourselves on the ProDA vignette:
https://rdrr.io/bioc/proDA/f/vignettes/Introduction.Rmd.

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
```

Important: make sure you have run 1_analysis_CPTAC_MSqRob.Rmd OR load `peptides.CPTAC` via:

```{r}
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC.RData"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

## 1. Data import and preprocessing

### 1.1. Import the data

1. **Important:** in the CPTAC dataset, some human UPS proteins are NOT contaminants as these proteins were spiked in on purpose! We only want to remove those contaminants that are not humane UPS proteins. We thus need to unmark these proteins as contaminant before preprocessing.

The following two proteins were manually indicated as not being contaminants in Perseus by removing their corresponding `"+"` symbol in the `"Potential contaminants"` column:

P06396ups;CON__Q3SX14

P02768ups;CON__P02768-1

We saved the new file as **`"proteinGroups_cont_curated.txt"`** (see 6_analysis_CPTAC_Perseus.Rmd).

```{r}
# Load the table into memory
dataset.CPTAC <- read.delim(
    paste0(wd, "/datasets/CPTAC/proteinGroups_cont_curated.txt"),
    stringsAsFactors = FALSE
)
```

### 1.2. Preprocess data

```{r}
# Remove reverse sequences, potential contaminants and proteins only identified by modified peptides
dataset.CPTAC <- dataset.CPTAC[rowSums(dataset.CPTAC[,c("Only.identified.by.site", "Reverse", "Potential.contaminant")] == "+") == 0,]

# Use a regex to select the intensity columns
intensity.colnames <- grep("^LFQ\\.intensity\\.", colnames(dataset.CPTAC), value=TRUE)
head(intensity.colnames)

# Create the intensity matrix
abundance.matrix <- as.matrix(dataset.CPTAC[, intensity.colnames])
# Adapt column and row dataset.CPTAC
colnames(abundance.matrix) <- sub("^LFQ\\.intensity\\.", "", intensity.colnames)
rownames(abundance.matrix) <- dataset.CPTAC$Protein.IDs

# Print some rows of the matrix
head(abundance.matrix)
```

```{r}
# Log2-transform
abundance.matrix[abundance.matrix == 0] <- NA
abundance.matrix <- log2(abundance.matrix)
```

### 1.3. Inspect the data

```{r}
barplot(colSums(is.na(abundance.matrix)),
        ylab = "# missing values",
        xlab = "Samples")

boxplot(abundance.matrix,
        ylab = "Intensity Distribution",
        xlab = "Samples")
```

## 2. ProDA analysis

The vignette suggests the more "conservative" median normalization, which ignores the missing values and transforms the values so that the median difference between the sample and average across all other samples is zero.

```{r}
normalized.abundance.matrix <- median_normalization(abundance.matrix)
colnames(normalized.abundance.matrix) <- peptides.CPTAC %>% sampleNames
```

```{r}
annotation <- data.frame(sample = peptides.CPTAC %>% sampleNames %>% as.character, condition = peptides.CPTAC %>% sampleNames %>% substr(2,3) %>% as.character, lab = c("LTQ-Orbitrap_86" %>% rep(3), "LTQ-OrbitrapO_65" %>% rep(3), "LTQ-OrbitrapW_56" %>% rep(3)) %>% rep(3) %>% as.character, stringsAsFactors = FALSE)

fit_ref6A <- proDA(normalized.abundance.matrix, design = ~ condition + lab, 
             col_data = annotation, reference_level = "6A")
fit_ref6B <- proDA(normalized.abundance.matrix, design = ~ condition + lab, 
             col_data = annotation, reference_level = "6B")
```

```{r}
res.CPTAC.ProDA_B_A <- test_diff(fit_ref6A, "condition6B")
res.CPTAC.ProDA_B_A$UPS <- grepl("ups", res.CPTAC.ProDA_B_A$name)
res.CPTAC.ProDA_B_A$contrast <- "condition6B-condition6A"
res.CPTAC.ProDA_B_A <- res.CPTAC.ProDA_B_A[,c("name", "UPS", "contrast", "diff", "se",  "t_statistic", "df", "pval", "adj_pval")]
colnames(res.CPTAC.ProDA_B_A) <- c("protein", "UPS", "contrast", "logFC", "se", "t", "df", "pval", "qval")

res.CPTAC.ProDA_C_A <- test_diff(fit_ref6A, "condition6C")
res.CPTAC.ProDA_C_A$UPS <- grepl("ups", res.CPTAC.ProDA_C_A$name)
res.CPTAC.ProDA_C_A$contrast <- "condition6C-condition6A"
res.CPTAC.ProDA_C_A <- res.CPTAC.ProDA_C_A[,c("name", "UPS", "contrast", "diff", "se",  "t_statistic", "df", "pval", "adj_pval")]
colnames(res.CPTAC.ProDA_C_A) <- c("protein", "UPS", "contrast", "logFC", "se", "t", "df", "pval", "qval")

res.CPTAC.ProDA_C_B <- test_diff(fit_ref6B, "condition6C")
res.CPTAC.ProDA_C_B$UPS <- grepl("ups", res.CPTAC.ProDA_C_B$name)
res.CPTAC.ProDA_C_B$contrast <- "condition6C-condition6B"
res.CPTAC.ProDA_C_B <- res.CPTAC.ProDA_C_B[,c("name", "UPS", "contrast", "diff", "se",  "t_statistic", "df", "pval", "adj_pval")]
colnames(res.CPTAC.ProDA_C_B) <- c("protein", "UPS", "contrast", "logFC", "se", "t", "df", "pval", "qval")

res.CPTAC.ProDA <- rbind(res.CPTAC.ProDA_B_A, res.CPTAC.ProDA_C_A, res.CPTAC.ProDA_C_B)
res.CPTAC.ProDA <- as_tibble(res.CPTAC.ProDA)
res.CPTAC.ProDA.full <- res.CPTAC.ProDA %>% arrange(pval)

res.CPTAC.ProDA.full

# Or load the results:
# load(paste0(wd, "/save_files_CPTAC/res_CPTAC_ProDA_full.RData"))
```
