---
title: "Count analysis for the CPTAC dataset"
author: "Ludger Goeminne"
date: "28/06/2019"
output: html_document
---

# The quasibinomial model for the CPTAC dataset

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
```

Important: make sure you have run 1_analysis_CPTAC_MSqRob.Rmd OR load `peptides.CPTAC2`, `prot.CPTAC.co` and `res.CPTAC.Base` via:

```{r}
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC2.RData"))
load(paste0(wd, "/save_files_CPTAC/prot_CPTAC_co.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_Base.RData"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

## Fit the generalized linear regression with quasi-binomial family and logit link function and inspect the results

```{r eval=TRUE}
form.co <- ~ condition + lab
contrasts <- contrast_helper(~ condition, peptides.CPTAC2, condition)

# res.CPTAC.co <- do_glm(formula = form.co, msnset = prot.CPTAC.co, group_var = protein, contrasts = contrasts, contFun = "contEst") ## 4 mins.
# OR: load the model:
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_co.RData"))

res.CPTAC.co.full <- res.CPTAC.Base %>% left_join(res.CPTAC.co$result)
res.CPTAC.co.full <- res.CPTAC.co.full %>% arrange(pvalue)

# Or load the MSqRob result object via:
# load(paste0(wd, "/save_files_CPTAC/res_CPTAC_co_full.RData"))

# Optional: remove res.CPTAC.co to free up memory space
rm(res.CPTAC.co)
gc()
```


