---
title: "Preparation of the spectral count data based on MaxQuant's evidence.txt file"
author: "Ludger Goeminne"
date: "26/02/2020"
output: html_document
---

# Preparation of the spectral count data based on MaxQuant's evidence.txt file

Set working directory and load all necessary libraries and functions.

```{r}
# If not installed, install and load the package "here", else: only load the package.
err <- try(library("here", character.only = TRUE), silent = TRUE)
if (class(err) == 'try-error') {
  install.packages("here", repos = "https://cloud.r-project.org")
  library("here", character.only = TRUE)
}

wd <- here()

# Optional: change the working directory
setwd(wd)

# Load all functions and libraries
source(paste0(wd, "/R/functions.r"))
```

Important: make sure you have run 1_analysis_CPTAC_MSqRob.Rmd OR load `peptides.CPTAC2`, `prot.CPTAC.co` and `res.CPTAC.Base` via:

```{r}
load(paste0(wd, "/save_files_CPTAC/peptides_CPTAC2.RData"))
load(paste0(wd, "/save_files_CPTAC/prot_CPTAC_co.RData"))
load(paste0(wd, "/save_files_CPTAC/res_CPTAC_Base.RData"))
```

Give session info for reproducibility.

```{r}
sessionInfo()
```

## 1. Prepare the spectral count data

Import the evidence.txt file.

```{r eval=TRUE}
# The evidence.txt file was with its 73.9 MB too big to be delivered on a GitHub server. However, you can load it in via the following .RData file:
load(paste0(wd, "/save_files_CPTAC/MSstats_infile.RData"))

# If you wish to recreate the evidence.txt file, simply write it out to a tab-delimited file again via:
# write.table(infile, file = "evidence.txt", quote = FALSE, sep = "\t", na = "", row.names = FALSE, col.names = TRUE)
# Then import it again the same way as we did via:
# infile <- read.table(paste0(wd,"/datasets/CPTAC/evidence.txt"), sep="\t", header=TRUE)
```

Preprocess the data.

```{r eval=TRUE}
#There are no intensities equal to 0, but there are some NA, remove them:
evidence <- subset(infile, !is.na(infile$Intensity))
evidence$feature <- paste(evidence$Modified.sequence, evidence$Charge, sep = "_")

evidencesplit <- split(evidence , f = evidence$Experiment)

all(rownames(exprs(prot.CPTAC.co)) %in% evidence$Proteins)

# Theoretical "maximum" number of spectral counts per protein: maximum number of counts per feature over all the samples
evidence2 <- evidence %>% group_by(feature, Proteins, Gene.names, Protein.names) %>%
  summarise(group_count = max(MS.MS.count))

# Then sum up the counts per protein
evidence3 <- evidence2 %>% group_by(Proteins, Gene.names, Protein.names) %>%
  summarise(group_count = sum(group_count))

fd1 <- fData(prot.CPTAC.co)
fd1$group_count <- NULL
dim(fd1)
fd <- left_join(fd1, evidence3, by = c("protein" = "Proteins", "gene.name" = "Gene.names", "protein.name" = "Protein.names"))
dim(fd)
# check
all(fd$group_count >= fd1$group_count)
# TRUE

# Some features are seen more than once...
evidencesplit[[1]][evidencesplit[[1]]$feature == "_AAATAAM(ox)TK__2",]

# Do as in the artMS package
# https://github.com/biodavidjm/artMS/blob/master/R/evidenceToSaintExpressFormat.R

evidencesplit2 <- lapply(evidencesplit, function(x){
  x <- x %>% group_by(feature, Proteins, Gene.names, Protein.names) %>%
    summarise(group_count = max(MS.MS.count))
  return(x)
})

evidencesplit3 <- lapply(evidencesplit2, function(x){
  x <- x %>% group_by(Proteins, Gene.names, Protein.names) %>%
    summarise(group_count = sum(group_count))
  return(x)
})

exprs <- matrix(0, nrow = nrow(fd), ncol = length(evidencesplit3), dimnames = dimnames(exprs(prot.CPTAC.co)))

for(i in 1:length(evidencesplit3)){
sample.evidence <- evidencesplit3[[i]]
sample.evidence <- sample.evidence[sample.evidence$Proteins %in% rownames(exprs),]

for(j in 1:nrow(sample.evidence)){
  sample.evidence2 <- sample.evidence[j, , drop = TRUE]
  exprs[rownames(exprs) == as.character(lol2[,"Proteins", drop = TRUE]), i] <- sample.evidence2[,"group_count", drop = TRUE]
}
}

pd <- pData(prot.CPTAC.co)[, c("condition", "lab")]
rownames(pd) <- sampleNames(prot.CPTAC.co)
rownames(fd) <- fd$protein

prot.CPTAC.spectral.co <- MSnSet(exprs, fData = AnnotatedDataFrame(fd), pData = AnnotatedDataFrame(pd))
# Or load the summarized object via:
# load(paste0(wd, "/save_files_CPTAC/prot_CPTAC_spectral_co.RData"))
```
